---
title: "Figure 3 ANCOM-BC2 — Cohort 2 differential abundance"
author: "Anna Petouhoff"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Runs **ANCOM-BC2** on Cohort 2 baseline stool samples (PPI vs no PPI) and
writes `../data/Cohort_2/ANCOM-BC2_response_baseline.RData`.

That file is already in the repository. Figures 3E–F load it directly, so this
notebook is needed only to *regenerate* the ANCOM-BC2 results, not to plot
3E–F.

Clinical variables used: `Patient Id` (join key) and `PPI`. Samples are
restricted to each patient's first stool collection within ±10 days of ICI
start (`Days` from microbiome metadata).

```{r libraries}
library(ANCOMBC)
library(phyloseq)
library(tidyverse)
```

```{r load}
load("../data/Cohort_2/counts-long.rds")
load("../data/Cohort_2/mitox_microbiome_metadata.rds")
load("../data/Cohort_2/clinical.rds")
```

```{r prepare}
# One baseline sample per patient, within 10 days of ICI start.
baseline <- mic.meta %>%
  group_by(`Patient Id`) %>%
  arrange(Days) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  filter(abs(Days) <= 10)

clin <- left_join(baseline, demographics, by = "Patient Id") %>%
  dplyr::filter(sample %in% counts.l$sample) %>%
  mutate(PPI = if_else(is.na(PPI), "No", as.character(PPI))) %>%
  dplyr::select(sample, PPI)

met <- clin %>%
  dplyr::filter(sample %in% counts.l$sample)

# Species-level counts (drop strain rows).
counts <- counts.l %>%
  dplyr::filter(sample %in% met$sample,
                grepl("s__", Taxonomy),
                !grepl("t__", Taxonomy)) %>%
  pivot_wider(names_from = "sample", values_from = "est.reads", values_fill = 0) %>%
  column_to_rownames("Taxonomy")

met <- met %>%
  arrange(factor(sample, levels = colnames(counts))) %>%
  column_to_rownames("sample")

all.equal(rownames(met), colnames(counts))
```

```{r metaphlan-to-phyloseq}
metaphlanToPhyloseq <- function(tax, metadat, simplenames = TRUE,
                                roundtointeger = FALSE, split = "|") {
  xnames <- rownames(tax)
  shortnames <- gsub(paste0(".+\\", split), "", xnames)
  if (simplenames) {
    rownames(tax) <- substr(shortnames, 4, nchar(shortnames))
  }
  tax <- as.matrix(tax)
  if (roundtointeger) {
    tax <- round(tax * 1e4)
  }
  x2 <- strsplit(xnames, split = split, fixed = TRUE)
  taxmat <- matrix(NA, ncol = max(sapply(x2, length)), nrow = length(x2))
  colnames(taxmat) <- c("Kingdom", "Phylum", "Class", "Order", "Family",
                        "Genus", "Species", "Strain")[1:ncol(taxmat)]
  rownames(taxmat) <- rownames(tax)
  for (i in 1:nrow(taxmat)) {
    taxmat[i, 1:length(x2[[i]])] <- x2[[i]]
  }
  taxmat <- gsub("[a-z]__", "", taxmat)
  taxmat <- phyloseq::tax_table(taxmat)
  otutab <- phyloseq::otu_table(tax, taxa_are_rows = TRUE)
  if (is.null(metadat)) {
    phyloseq::phyloseq(taxmat, otutab)
  } else {
    phyloseq::phyloseq(taxmat, otutab, phyloseq::sample_data(metadat))
  }
}

phylo <- metaphlanToPhyloseq(counts, met)
phylo
```

```{r run-ancombc2}
run_ancombc2 <- function(level, grp) {
  out <- ANCOMBC::ancombc2(
    data = phylo,
    tax_level = level,
    fix_formula = grp,
    group = grp,
    struc_zero = TRUE,
    s0_perc = 0,
    prv_cut = 0,
    p_adj_method = "BH"
  )
  list(res = out$res, struct0 = out$zero_ind, feature_tab = out$feature_table)
}

tax_levels <- c("Species", "Genus", "Family", "Order", "Class", "Phylum", "Kingdom")
res_list <- lapply(tax_levels, function(level) run_ancombc2(level, "PPI"))
res.recist <- do.call(rbind, lapply(res_list, function(x) x$res))
struct0.recist <- do.call(rbind, lapply(res_list, function(x) x[[2]]))
feature.recist <- do.call(rbind, lapply(res_list, function(x) x[[3]]))

save(res.recist, struct0.recist, feature.recist,
     file = "../data/Cohort_2/ANCOM-BC2_response_baseline.RData")
```

```{r check-duplicates}
# Duplicate taxa names are unclassified features at multiple ranks.
duplicates <- duplicated(res.recist$taxon)
res.recist[duplicates | duplicated(res.recist$taxon, fromLast = TRUE), ]
```
