---
title: "Figure 4 — Cohort 3 microbiome, CyTOF, and correlation"
author: "Anna Petouhoff"
date: "2025-03-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Produces:

- **Figure 4B**: phylum stacked bars for baseline stool samples, by PPI use
- **Figure 4C**: CyTOF immune-cell abundance heatmap at ICI start
- **Figure 4D**: Spearman correlation of *Enterocloster clostridioformis*
  relative abundance with terminal-effector CD4 T cells

Cohort 3 is the n = 8 NSCLC/RCC patients with both a baseline microbiome
sample and CyTOF. Sample names in the MetaPhlAn table are study IDs
(`P026_d0`, …) with days relative to ICI start; they do not contain calendar
dates. The only clinical variable used here is PPI use
(`../data/Cohort_3/clinical.csv`).

```{r libraries}
library(tidyverse)
library(broom)
library(ggplot2)
library(pheatmap)
library(viridis)
```

```{r load}
micro <- read.csv("../data/Cohort_3/2023-12-07_mpa-aggregate.csv", stringsAsFactors = FALSE)
cytof <- readRDS("../data/Cohort_3/cytof_normalized.RDS")
clin3 <- read.csv("../data/Cohort_3/clinical.csv", stringsAsFactors = FALSE)
```

```{r ids}
# Map recoded stool sample IDs to the CyTOF Patient ID scheme ("26 _ 0").
# d0 = baseline; other suffixes are follow-up samples (excluded from figures).
sample_map <- tibble(
  sample = c("P026_d0", "P028_d0", "P029_d0", "P031_d0", "P032_d0",
             "P033_d0", "P034_d0", "P041_d0",
             "P031_d28", "P032_d22", "P033_d21", "P034_d23"),
  patient = c(26, 28, 29, 31, 32, 33, 34, 41, 31, 32, 33, 34),
  day = c(0, 0, 0, 0, 0, 0, 0, 0, 28, 22, 21, 23)
) %>%
  mutate(`Patient ID` = paste(patient, day, sep = " _ ")) %>%
  left_join(clin3, by = "patient")

# Baseline stool samples used in the published panels.
baseline_map <- sample_map %>% filter(day == 0)
```

# Species-level relative abundance, baseline samples only

```{r species}
species <- micro %>%
  filter(grepl("s_", Taxonomy), !grepl("t_", Taxonomy)) %>%
  inner_join(baseline_map, by = "sample") %>%
  mutate(Taxonomy = sub(".*s__", "", Taxonomy))

species_wide <- species %>%
  select(`Patient ID`, Taxonomy, RelAbun, PPI) %>%
  pivot_wider(names_from = Taxonomy, values_from = RelAbun, values_fill = 0)
```

# Merge baseline CyTOF with microbiome

```{r merge}
cytof_baseline <- cytof %>%
  filter(`Patient ID` %in% baseline_map$`Patient ID`)

merged <- inner_join(species_wide, cytof_baseline, by = "Patient ID")

# Drop immune-cell columns with missing or non-finite values (markers not
# acquired for this batch).
immune_cols <- setdiff(names(cytof_baseline), c("Patient ID", "Time"))
keep_immune <- immune_cols[vapply(merged[immune_cols], function(x) {
  all(is.finite(x)) && sum(!is.na(x)) > 1
}, logical(1))]
cytof_mat <- merged[, keep_immune, drop = FALSE]
```

# Figure 4D: Enterocloster clostridioformis vs TE CD4

```{r figure-4d}
ggplot(merged, aes(x = `TE CD4`, y = Enterocloster_clostridioformis)) +
  geom_point() +
  stat_smooth(method = "lm", aes(group = 1)) +
  labs(
    x = "TE CD4 Count",
    y = expression(atop("Relative Abundance of", italic("Enterocloster clostridioformis")))
  ) +
  theme_minimal(base_size = 8)

ggsave("../figures/Figure4D.svg", device = "svg", height = 1.75, width = 5)
```

# Exploratory Spearman correlations (text; FDR-adjusted)

```{r correlations}
# Two PPI-associated species carried into the immune-cell screen.
interest <- merged %>%
  select(Clostridium_sp_AT4, Enterocloster_clostridioformis) %>%
  mutate(across(everything(), ~ replace(., is.na(.), 0)))

set.seed(9)
corr.results <- lapply(colnames(interest), function(x) {
  lapply(keep_immune, function(y) {
    cor.test(interest[[x]], merged[[y]], method = "spearman", exact = FALSE) %>%
      tidy() %>%
      mutate(microbe = x, cytof = y)
  }) %>% bind_rows()
}) %>%
  bind_rows() %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr"))
```

# Figure 4B: phylum stacked bars

```{r figure-4b}
phyl.ra <- micro %>%
  filter(grepl("p__", Taxonomy), !grepl("c__", Taxonomy)) %>%
  inner_join(baseline_map, by = "sample")

top.phyl <- phyl.ra %>%
  group_by(Taxonomy) %>%
  summarise(med.abun = median(RelAbun, na.rm = TRUE)) %>%
  arrange(desc(med.abun)) %>%
  filter(med.abun > 0.85)

phyl.ra <- phyl.ra %>%
  mutate(Phylum = ifelse(Taxonomy %in% top.phyl$Taxonomy,
                         gsub("p__", "", Taxonomy), "Other"))
phyl.ra$Phylum <- sub("^[^|]*\\|", "", phyl.ra$Phylum)

phyl.ra <- phyl.ra %>%
  group_by(sample) %>%
  mutate(Firmicutes_Abundance = sum(RelAbun[Phylum == "Firmicutes"], na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(sample = fct_reorder(sample, Firmicutes_Abundance)) %>%
  mutate(Phylum = factor(Phylum, levels = c(
    "Firmicutes",
    setdiff(unique(Phylum), c("Firmicutes", "Other")),
    "Other"
  )))

bar_colors <- c("cyan4", "darkgoldenrod2", "darkgreen", "salmon2",
                "cornflowerblue", "deeppink4", "#999933")

phyl.ra %>%
  ggplot(aes(y = `Patient ID`, x = RelAbun, fill = Phylum)) +
  geom_bar(position = "fill", stat = "identity", width = 0.7,
           color = "white", linewidth = 0.1) +
  scale_fill_manual(values = bar_colors) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(y = "Patients", x = "Relative Abundance") +
  theme_minimal(base_size = 9) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.title = element_blank(),
    legend.text = element_text(size = 6),
    legend.key.size = unit(0.3, "cm"),
    legend.spacing.x = unit(0.1, "cm"),
    strip.text.y = element_text(angle = 0, face = "bold"),
    strip.background = element_rect(fill = "gray95", color = NA),
    plot.margin = margin(t = 5, r = 10, b = 5, l = 5)
  ) +
  guides(fill = guide_legend(ncol = 4, byrow = TRUE)) +
  facet_grid(PPI ~ ., scales = "free_y", space = "free_y")

ggsave("../figures/Figure4B.svg", device = "svg", height = 3, width = 3.25)
```

# Figure 4C: CyTOF heatmap, columns ordered by PPI status

```{r figure-4c}
immune_matrix <- t(as.matrix(cytof_mat))
colnames(immune_matrix) <- merged$`Patient ID`

ppi_annotation <- data.frame(
  PPI = merged$PPI,
  row.names = merged$`Patient ID`
)
ordered_columns <- order(merged$PPI)
ordered_immune_matrix <- immune_matrix[, ordered_columns, drop = FALSE]
ordered_ppi_annotation <- ppi_annotation[ordered_columns, , drop = FALSE]

annotation_colors <- list(PPI = c("Y" = "lightsalmon2", "N" = "slateblue2"))

empty_names <- rep("", ncol(ordered_immune_matrix))
empty_names[round(length(empty_names) / 2)] <- "Patient Samples"

svg("../figures/Figure4C.svg", width = 4.25, height = 5.75)
pheatmap(
  ordered_immune_matrix,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  color = viridis(100),
  fontsize_row = 9,
  labels_col = empty_names,
  angle_col = 0,
  fontsize_col = 10,
  annotation_col = ordered_ppi_annotation,
  annotation_colors = annotation_colors
)
dev.off()
```
