---
title: "Figure 2E — PPI vs H2B sliding-window hazard ratios"
author: "Anna Petouhoff"
date: "2025-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Produces **Figure 2E**: Cox hazard ratios for PPI and H2B exposure in rolling
30-day windows relative to ICI start (day 0).

Clinical variables used: `days`, `vitalstatus`, and the relative-day coverage
strings `PPI.days.to.iostart` and `H2B.days.to.iostart` (days on drug relative
to ICI start, not calendar dates).

Cleaned Cohort 1 object: `../data/cohort1-cleaned-clinical.RDS`

```{r libraries}
library(tidyverse)
library(survival)
library(glue)
```

```{r load}
db <- readRDS("../data/cohort1-cleaned-clinical.RDS") %>%
  select(days, vitalstatus, PPI.days.to.iostart, H2B.days.to.iostart)
```

```{r window-helper}
# Return 1 if the comma-separated relative-day string overlaps [per[1], per[2]].
check.window <- function(vect, per) {
  wind <- per[1]:per[2]
  wind <- glue_collapse(as.character(wind), sep = ",|,")
  wind <- paste(",", wind, ",", sep = "")
  ifelse(grepl(pattern = wind, x = vect), 1, 0)
}

# ±28-day exposure at ICI start, used to define the unexposed reference group.
db <- db %>%
  mutate(ppi.28.28 = check.window(paste(",", PPI.days.to.iostart, ",", sep = ""), c(-28, 28))) %>%
  mutate(h2b.28.28 = check.window(paste(",", H2B.days.to.iostart, ",", sep = ""), c(-28, 28)))
```

# PPI: 30-day windows from day −180 to +151

```{r ppi-windows, message=FALSE, error=FALSE, warning=FALSE}
ppifloattime <- db %>%
  select(days, vitalstatus, PPI.days.to.iostart) %>%
  mutate(PPI.days.to.iostart = paste(",", PPI.days.to.iostart, ",", sep = ""))
ppi.zero.cohort <- db %>%
  select(days, vitalstatus, ppi.28.28) %>%
  filter(ppi.28.28 == 0) %>%
  mutate(tmp = ppi.28.28) %>%
  select(days, vitalstatus, tmp)
leftanchor <- c(-180:151)
ppiwindresults <- list()
i <- 1
ppifloattime$tmp2 <- check.window(ppifloattime$PPI.days.to.iostart, c(leftanchor[i], leftanchor[i] + 29))

for (i in 1:length(leftanchor)) {
  j <- leftanchor[i]
  ppifloattime$tmp <- check.window(ppifloattime$PPI.days.to.iostart, c(j, j + 29))

  # Keep patients whose exposure status is stable or newly starts in this window.
  ppifloattime <- ppifloattime %>%
    mutate(tmp3 = tmp2 - tmp) %>%
    filter(tmp3 == 0 | tmp3 == -1) %>%
    mutate(tmp2 = tmp)

  ppifloatforsurv <- ppifloattime %>%
    filter(tmp == 1) %>%
    bind_rows(ppi.zero.cohort)

  try({
    survtemp <- coxph(Surv(days, vitalstatus) ~ tmp, data = ppifloatforsurv) %>%
      summary()
    pvals <- survtemp$coefficients %>%
      as.data.frame() %>%
      rownames_to_column(var = "term") %>%
      mutate(p.value = `Pr(>|z|)`) %>%
      select(term, p.value)
    conf.int <- survtemp$conf.int %>%
      as.data.frame() %>%
      rownames_to_column(var = "term") %>%
      mutate(hazard.ratio = `exp(coef)`,
             confint.low = `lower .95`,
             confint.high = `upper .95`,
             leftanchor = j) %>%
      select(term, hazard.ratio, confint.low, confint.high, leftanchor) %>%
      left_join(pvals)
    ppiwindresults[[i]] <- conf.int
  })
}

ppipvaltime <- bind_rows(ppiwindresults)
# Plot at the window midpoint.
ppipvaltime$leftanchor <- ppipvaltime$leftanchor + 15
```

# H2B: same window series

```{r h2b-windows, message=FALSE, error=FALSE, warning=FALSE}
h2bfloattime <- db %>%
  select(days, vitalstatus, H2B.days.to.iostart) %>%
  mutate(H2B.days.to.iostart = paste(",", H2B.days.to.iostart, ",", sep = ""))
h2b.zero.cohort <- db %>%
  select(days, vitalstatus, h2b.28.28) %>%
  filter(h2b.28.28 == 0) %>%
  mutate(tmp = h2b.28.28) %>%
  select(days, vitalstatus, tmp)
leftanchor <- c(-180:151)
h2bwindresults <- list()
i <- 1
h2bfloattime$tmp2 <- check.window(h2bfloattime$H2B.days.to.iostart, c(leftanchor[i], leftanchor[i] + 29))

for (i in 1:length(leftanchor)) {
  j <- leftanchor[i]
  h2bfloattime$tmp <- check.window(h2bfloattime$H2B.days.to.iostart, c(j, j + 29))

  h2bfloattime <- h2bfloattime %>%
    mutate(tmp3 = tmp2 - tmp) %>%
    filter(tmp3 == 0 | tmp3 == -1) %>%
    mutate(tmp2 = tmp)

  h2bfloatforsurv <- h2bfloattime %>%
    filter(tmp == 1) %>%
    bind_rows(h2b.zero.cohort)

  try({
    survtemp <- coxph(Surv(days, vitalstatus) ~ tmp, data = h2bfloatforsurv) %>%
      summary()
    pvals <- survtemp$coefficients %>%
      as.data.frame() %>%
      rownames_to_column(var = "term") %>%
      mutate(p.value = `Pr(>|z|)`) %>%
      select(term, p.value)
    conf.int <- survtemp$conf.int %>%
      as.data.frame() %>%
      rownames_to_column(var = "term") %>%
      mutate(hazard.ratio = `exp(coef)`,
             confint.low = `lower .95`,
             confint.high = `upper .95`,
             leftanchor = j) %>%
      select(term, hazard.ratio, confint.low, confint.high, leftanchor) %>%
      left_join(pvals)
    h2bwindresults[[i]] <- conf.int
  })
}

h2bpvaltime <- bind_rows(h2bwindresults)
h2bpvaltime$leftanchor <- h2bpvaltime$leftanchor + 15
```

```{r figure-2e, fig.height=8, fig.width=10}
ppipvaltime$source <- "PPI"
h2bpvaltime$source <- "H2B"
allpvaltime <- bind_rows(ppipvaltime, h2bpvaltime)
ggplot(data = allpvaltime, aes(x = leftanchor, y = hazard.ratio)) +
  geom_line(aes(color = source)) +
  geom_ribbon(aes(ymin = confint.low, ymax = confint.high, fill = source), alpha = 0.5) +
  geom_hline(yintercept = 1) +
  scale_color_manual(aesthetics = c("fill", "color"), breaks = c("PPI", "H2B"),
                     values = c("darkred", "darkgreen"), name = "Drug") +
  labs(title = "", x = "Days relative to ICI start", y = "Hazard Ratio") +
  theme_bw() +
  theme(axis.text = element_text(size = 8), axis.title = element_text(size = 8))

ggsave("../figures/Figure2E.svg", device = "svg", width = 6, height = 2)
```
