---
title: "Figures 1D–F and 2A–D — Cohort 1 Kaplan–Meier curves"
author: "Anna Petouhoff"
date: "2025-02-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Produces **Figures 1D–F and 2A–D**: overall survival by PPI use in Cohort 1,
overall and stratified by line of therapy, prior chemotherapy, and timing of
chemotherapy relative to ICI start.

Clinical variables used: `days`, `vitalstatus`, `ppi`, `line_of_therapy`,
`prior_chemo`, `chemo_greater_one_year`. No identifiers.

Cleaned Cohort 1 object: `../data/cohort1-cleaned-clinical.RDS`

Univariable Cox HRs reported in the text are computed in
`Figures 1 and 2 Univariable Cox PH.Rmd`.

```{r libraries}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(survival)
library(survminer)
library(patchwork)
```

```{r load}
CoRR <- readRDS("../data/cohort1-cleaned-clinical.RDS") %>%
  select(days, vitalstatus, ppi, line_of_therapy, prior_chemo, chemo_greater_one_year)
```

```{r km-helper}
# Shared Kaplan–Meier styling used for every panel in Figures 1 and 2.
plot_ppi_km <- function(fit, title) {
  p <- ggsurvplot(
    fit,
    pval = TRUE,
    risk.table = TRUE,
    title = title,
    legend.title = "",
    ylab = "Survival Probability",
    xlab = "Time (days)",
    legend.labs = c("Not Prescribed", "Prescribed PPIs"),
    palette = c("slateblue2", "lightsalmon2"),
    legend = c(0.8, 0.8),
    font.main = 9,
    font.x = 8,
    font.y = 8,
    font.tickslab = 7,
    font.legend = 7,
    pval.size = 3.5,
    pval.coord = c(100, 0.15),
    risk.table.fontsize = 2.8,
    tables.theme = theme_survminer()
  )
  p$table$theme$axis.text.y <- element_text(size = 7, color = "black")
  p$table$theme$plot.title <- element_blank()
  p$table <- p$table +
    theme(
      axis.text.x = element_text(size = 7),
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      axis.ticks.length = unit(0.05, "cm"),
      plot.margin = margin(t = -15, r = 5, b = 5, l = 5)
    )
  p$plot <- p$plot +
    theme(
      axis.title.y = element_text(vjust = -15, size = 8),
      plot.margin = margin(t = 5, r = 5, b = -15, l = 5)
    )
  p$plot / p$table + plot_layout(heights = c(4, 1))
}

save_km <- function(plot, file) {
  ggsave(file, plot = plot, device = "svg", width = 3.5, height = 2.5)
}
```

# Figure 1D: all cancers

```{r figure-1d}
all_surv <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR)
save_km(plot_ppi_km(all_surv, "OS For All Cancers"), "../figures/Figure1D.svg")
```

# Figure 1E: first-line ICI

```{r figure-1e}
first_line <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$line_of_therapy == 1, ])
save_km(plot_ppi_km(first_line, "OS For First Line ICIs"), "../figures/Figure1E.svg")
```

# Figure 1F: second-line ICI and beyond

```{r figure-1f}
second_line <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$line_of_therapy != 1, ])
save_km(plot_ppi_km(second_line, "OS For Second Line ICIs"), "../figures/Figure1F.svg")
```

# Figure 2A: no prior chemotherapy

```{r figure-2a}
no_chemo <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$prior_chemo == "no", ])
save_km(plot_ppi_km(no_chemo, "OS For No Prior Chemotherapy"), "../figures/Figure2A.svg")
```

# Figure 2B: prior chemotherapy

```{r figure-2b}
chemo <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$prior_chemo == "yes", ])
save_km(plot_ppi_km(chemo, "OS For Prior Chemotherapy"), "../figures/Figure2B.svg")
```

# Figure 2C: last chemotherapy ≥ 1 year before ICI

```{r figure-2c}
more_than_one_year <- survfit(Surv(days, vitalstatus) ~ ppi,
                              data = CoRR[CoRR$chemo_greater_one_year == 1, ])
save_km(plot_ppi_km(more_than_one_year, "OS For Chemotherapy Greater than One Year"),
        "../figures/Figure2C.svg")
```

# Figure 2D: last chemotherapy < 1 year before ICI

```{r figure-2d}
less_than_one_year <- survfit(Surv(days, vitalstatus) ~ ppi,
                              data = CoRR[CoRR$chemo_greater_one_year == 0, ])
save_km(plot_ppi_km(less_than_one_year, "OS For Chemotherapy Less than One Year"),
        "../figures/Figure2D.svg")
```
