---
title: "Figures 2F and 2G — multivariable Cox forest plots"
author: "Anna Petouhoff"
date: "2026-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Produces **Figures 2F and 2G**: hazard ratios for PPI (2F) or H2B (2G) from
three bivariate Cox models, each adjusting for one of: antibiotic use at ICI
start, cancer staging, or Charlson Comorbidity Index.

Clinical variables used: `ppi`, `h2b_iostart`, `abx_iostart`, `staging`,
`cci_number`, `days`, `vitalstatus`.

Cleaned Cohort 1 object: `../data/cohort1-cleaned-clinical.RDS`

```{r libraries}
library(survival)
library(dplyr)
library(tidyverse)
library(broom)
```

```{r load}
CoRR <- readRDS("../data/cohort1-cleaned-clinical.RDS") %>%
  select(ppi, h2b_iostart, abx_iostart, staging, cci_number, days, vitalstatus)
```

# Figure 2F: PPI, each model adjusting for one covariate

```{r ppi-models}
ppi_abx <- coxph(Surv(days, vitalstatus) ~ ppi + abx_iostart, data = CoRR)
ppi_staging <- coxph(Surv(days, vitalstatus) ~ ppi + staging, data = CoRR)
ppi_cci <- coxph(Surv(days, vitalstatus) ~ ppi + cci_number, data = CoRR)

res1 <- tidy(ppi_abx, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "PPI, controlling for Abx")
res2 <- tidy(ppi_staging, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "PPI, controlling for Staging")
res3 <- tidy(ppi_cci, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "PPI, controlling for CCI")

ppi_only <- bind_rows(res1, res2, res3) %>%
  filter(term == "ppi") %>%
  mutate(p_label = ifelse(p.value < 0.001, "p < 0.001", paste0("p = ", round(p.value, 3))))

ppi_fp <- ggplot(ppi_only, aes(x = estimate, y = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "slateblue") +
  geom_point(size = 3.5, color = "slateblue4") +
  geom_text(aes(label = p_label, x = conf.high), hjust = -0.1, vjust = 0.5, size = 4) +
  scale_x_log10(expand = expansion(mult = c(0.1, 0.6))) +
  labs(x = "Hazard Ratio (Log Scale)", y = "",
       title = "PPI Impact on Survival Across Bivariate Models") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11, color = "black"),
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    plot.margin = margin(t = 10, r = 20, b = 10, l = 10),
    panel.grid.minor = element_blank()
  )
ppi_fp
ggsave("../figures/Figure2F.svg", plot = ppi_fp, device = "svg", width = 6, height = 3)
```

# Figure 2G: H2B, same three adjustment models

```{r h2b-models}
h2b_abx <- coxph(Surv(days, vitalstatus) ~ h2b_iostart + abx_iostart, data = CoRR)
h2b_staging <- coxph(Surv(days, vitalstatus) ~ h2b_iostart + staging, data = CoRR)
h2b_cci <- coxph(Surv(days, vitalstatus) ~ h2b_iostart + cci_number, data = CoRR)

res4 <- tidy(h2b_abx, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "H2B, controlling for Abx")
res5 <- tidy(h2b_staging, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "H2B, controlling for Staging")
res6 <- tidy(h2b_cci, exponentiate = TRUE, conf.int = TRUE) %>% mutate(model = "H2B, controlling for CCI")

h2b_only <- bind_rows(res4, res5, res6) %>%
  filter(term == "h2b_iostart") %>%
  mutate(p_label = ifelse(p.value < 0.001, "p < 0.001", paste0("p = ", round(p.value, 3))))

h2b_fp <- ggplot(h2b_only, aes(x = estimate, y = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "grey50") +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "slateblue") +
  geom_point(size = 3.5, color = "slateblue4") +
  geom_text(aes(label = p_label, x = conf.high), hjust = -0.2, vjust = -0.5, size = 3.5) +
  scale_x_log10(expand = expansion(mult = c(0.1, 0.4))) +
  labs(x = "Hazard Ratio (Log Scale)", y = "",
       title = "H2B Impact on Survival Across Bivariate Models") +
  theme_minimal() +
  theme(
    axis.title.x = element_text(size = 12, margin = margin(t = 10)),
    axis.text.x = element_text(size = 11),
    axis.text.y = element_text(size = 11, color = "black"),
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    panel.grid.minor = element_blank()
  )
h2b_fp
ggsave("../figures/Figure2G.svg", plot = h2b_fp, device = "svg", width = 6, height = 3)
```
