---
title: "Figure 1C — concomitant medication survival heatmap"
author: "Anna Petouhoff"
date: "2025-04-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Produces **Figure 1C**: heatmap of log-rank p-values for overall survival by
concomitant medication, overall and stratified by cancer type. Tile color is
the Kaplan–Meier p-value; `+` / `-` mark the direction of the univariable Cox HR.

Clinical variables used: `vitalstatus`, `days`, medication flags (`ppi`,
`statin`, `h2b_iostart`, `steroid_iostart`, `nsaid_iostart`), cancer type
(`cancer.name`), and relative antibiotic-day strings (`ABX.days.to.iostart`)
to define a ±28-day window around ICI start. No identifiers or calendar dates.

Cleaned Cohort 1 object: `../data/cohort1-cleaned-clinical.RDS`

```{r libraries}
library(tidyverse)
library(survival)
library(survminer)
library(glue)
library(forcats)
```

```{r load}
db <- readRDS("../data/cohort1-cleaned-clinical.RDS") %>%
  select(
    vitalstatus, days, ppi, statin, h2b_iostart, steroid_iostart, nsaid_iostart,
    cancer.name, ABX.days.to.iostart
  ) %>%
  mutate(
    cancer.aggregated = ifelse(
      cancer.name %in% c(
        "non-small cell lung cancer", "melanoma",
        "renal cell carcinoma", "head and neck carcinoma"
      ),
      cancer.name,
      "Other"
    )
  )
```

```{r abx-window}
# Flag antibiotic coverage in a ±28-day window around ICI start (day 0).
check.window <- function(vect, per) {
  wind <- per[1]:per[2]
  wind <- glue_collapse(as.character(wind), sep = ",|,")
  wind <- paste(",", wind, ",", sep = "")
  ifelse(grepl(pattern = wind, x = vect), 1, 0)
}

db <- db %>%
  mutate(abx_start = check.window(ABX.days.to.iostart, c(-28, 28)))

# Medication columns plotted on the heatmap (x-axis order matches scale_x labels).
meds_iostart <- c("h2b_iostart", "nsaid_iostart", "ppi", "statin", "steroid_iostart", "abx_start")
allcancers <- c("Other", "melanoma", "non-small cell lung cancer",
                "renal cell carcinoma", "head and neck carcinoma")
```

# Univariable OS models in the full cohort (heatmap "All" row)

```{r all-cohort-models}
forms <- list()
for (var in meds_iostart) {
  forms[[var]] <- as.formula(paste("Surv(days, vitalstatus) ~ ", var, sep = ""))
}
sfits <- surv_fit(formula = forms, data = db)
names(sfits) <- meds_iostart

almeds.pval <- lapply(sfits, function(x) surv_pvalue(x)[1, 2])
names(almeds.pval) <- names(sfits)
almeds.hr <- lapply(forms, function(x) coefficients(summary(coxph(formula = x, data = db)))[, 2])
almeds.pval <- bind_rows(almeds.pval)
almeds.hr <- bind_rows(almeds.hr)
almeds.dat <- as.data.frame(bind_rows(almeds.pval, almeds.hr))
rownames(almeds.dat) <- c("p.value", "hazrat")
almeds.dat <- almeds.dat %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "meds") %>%
  mutate(labcol = ifelse(hazrat < 1, "red", "black"),
         cancer = "All") %>%
  select(cancer, meds, p.value, labcol)
```

# Repeat models within each cancer group

```{r by-cancer-models, message=FALSE, warning=FALSE, error=FALSE}
sfits <- list()
pvalbycancer <- list()
pvals <- list()
labcol <- list()
labs <- list()
for (canc in allcancers) {
  tmp <- subset(db, db$cancer.aggregated == canc)
  for (s in meds_iostart) {
    f <- as.formula(paste("Surv(days, vitalstatus) ~ ", s, sep = ""))
    try({
      sfits[[s]] <- surv_fit(formula = f, data = tmp)
      pvals[[s]] <- surv_pvalue(sfits[[s]])[1, 2]
      x <- coefficients(summary(coxph(formula = f, data = tmp)))[, 2]
      labs[[s]] <- ifelse(x < 1, "red", "black")
    })
  }
  pvalbycancer[[canc]] <- pvals
  labcol[[canc]] <- labs
  sfits <- list()
  pvals <- list()
  labs <- list()
}
allcancpvals <- bind_rows(pvalbycancer)
rownames(allcancpvals) <- names(pvalbycancer)
labcol <- bind_rows(labcol)
rownames(labcol) <- names(pvalbycancer)
```

```{r assemble-heatmap-data}
labcol <- labcol %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "meds", value = "labcol", -cancer) %>%
  mutate(labcol = ifelse(is.na(labcol), "black", labcol))
heat.data <- allcancpvals %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "meds", value = "p.value", -cancer) %>%
  full_join(labcol) %>%
  full_join(almeds.dat)
heat.data <- bind_rows(almeds.dat, heat.data)
```

```{r figure-1c, fig.height=7, fig.width=6}
heat.plot <- heat.data %>%
  ggplot(aes(meds, fct_relevel(cancer, "Other", "melanoma", "non-small cell lung cancer",
                               "renal cell carcinoma", "head and neck carcinoma", "All"))) +
  geom_tile(aes(fill = p.value)) +
  labs(x = "", y = "") +
  geom_point(aes(shape = labcol), size = 3) +
  scale_fill_distiller(
    guide = "legend",
    breaks = c(0.000000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1),
    limits = c(0, 1),
    values = c(0.000000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1),
    type = "div", palette = "RdBu", direction = 1,
    labels = c("<0.01", "0.01", "0.02", "0.03", "0.04", "0.05", "0.1", "0.2", "0.5", "1")
  ) +
  scale_shape_manual(breaks = c("black", "red"), values = c("-", "+"),
                     labels = c("HR>1", "HR<1"), name = "Direction") +
  theme_classic(base_size = 9) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title = element_blank(),
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 8),
    legend.margin = margin(0, 0, 0, 0),
    legend.spacing.y = unit(0.1, "cm")
  ) +
  scale_x_discrete(
    breaks = meds_iostart,
    labels = c("H2B", "NSAIDs", "PPI", "Statins", "CS", "ABx")
  ) +
  scale_y_discrete(
    breaks = c("Other", "melanoma", "non-small cell lung cancer",
               "renal cell carcinoma", "head and neck carcinoma", "All"),
    labels = c("Other", "Melanoma", "Non-small cell\nlung cancer",
               "Renal cell\ncarcinoma", "Head and\nneck \ncarcinoma", "All")
  )
heat.plot

ggsave("../figures/Figure1C.svg", device = "svg", width = 3, height = 3.75)
```
