---
title: "Cohort 1 clinical processing (de-identification)"
author: "Anna Petouhoff"
date: "2025-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Directory of this notebook, used to find 00-paths.R and to write the RDS.
this_dir <- if (!is.null(knitr::current_input())) {
  dirname(normalizePath(knitr::current_input()))
} else if (file.exists("Processing.Rmd")) {
  getwd()
} else if (file.exists("manuscript/scripts/Processing.Rmd")) {
  file.path(getwd(), "manuscript/scripts")
} else {
  getwd()
}

paths_file <- file.path(this_dir, "00-paths.R")
if (!file.exists(paths_file)) {
  stop(
    "Missing 00-paths.R. Copy manuscript/scripts/00-paths.R.example to ",
    "00-paths.R and set Tdrive to the ppiio folder on the lab share."
  )
}
source(paths_file)
if (!exists("Tdrive") || !nzchar(Tdrive)) {
  stop("00-paths.R must define Tdrive as the ppiio project folder on the lab share.")
}
```

This notebook builds the **de-identified Cohort 1 clinical object** used by all
Figure 1, Figure 2, Table 1, and Supplementary Figure 3 scripts.

**PHI handling.** Raw REDCap exports and medication tables live on the lab
share (root set as `Tdrive` in gitignored `00-paths.R`; see
`00-paths.R.example`). Those files contain identifiers (MRN, date of birth,
calendar dates of treatment and death). Those columns are used only to derive
age, follow-up time, and medication timing relative to ICI start. The object
written at the end contains **no identifiers and no calendar dates** — only
the clinical variables that appear in the published analyses.

**Output (cleaned clinical object, committed with the repo):**
`manuscript/data/cohort1-cleaned-clinical.RDS`

This notebook must be run on a machine with the lab share in order to *create*
that RDS. Figure and table scripts read the RDS from `manuscript/data/` and
do not need the raw files.

```{r libraries}
library(dplyr)
library(tidyr)
library(tidyverse)
library(readxl)
```

# Raw inputs (PHI; lab share only)

Relative paths below are joined to `Tdrive` from `00-paths.R`. Do not put a
leading slash on the second `file.path()` argument; that would discard `Tdrive`
on some systems.

```{r read-raw}
CoRR <- read.csv(file.path(Tdrive, "data/raw/CoRR4684/2025-02-17/CoRR4684ARetrospecti_DATA_2025-02-17_2009.csv"))
meds <- read_excel(file.path(Tdrive, "data/raw/IW-meds/PLATELET_Variable_update.xlsx"), sheet = 2, skip = 1)
```

# Recode keys for cancer type and immunotherapy class

REDCap stores cancer type and ICI regimen as integers. Map them to the labels
used in Table 1 and Figure 1C.

```{r recode-keys}
cancer.copy <- c("1 non-small cell lung cancer
2 small cell lung cancer
3 melanoma
4 renal cell carcinoma
5 head and neck carcinoma
6 merkel cell carcinoma
7 Hodgkin lymphoma
8 breast cancer
9 colon cancer
10 pancreatic cancer
11 Sarcoma
12 Prostate
13 bladder cancer
14 AML
15 ALL
17 Other") %>%
  strsplit(., split = "\n") %>%
  unlist

cancer.key <-
  data.frame(
    cancer = cancer.copy %>%
      gsub("(\\d+).*", "\\1", .) %>%
      as.integer(),
    cancer.name = cancer.copy %>%
      gsub("\\d+ (.*)", "\\1", .),
    stringsAsFactors = FALSE
  )

ionames <- data.frame(
  immunotherapy = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
  Immunotherapy = c(
    "Nivolumab", "Pembrolizumab",
    "Atezolizumab", "Ipilimumab",
    "Nivolumab + Ipilimumab",
    "Durvalumab + Tremelimumab",
    "Tremelimumab", "Nivolumab + Chemotherapy",
    "Durvalumab", "Other"
  ),
  stringsAsFactors = FALSE
)
```

# Medication class lists (PPI, H2B, antibiotics)

Used to flag exposure and to compute days-on-drug relative to ICI start
(Figures 1C, 2E, 2F).

```{r med-class-helpers}
pull_class_assignments <- function() {
  h2b <- read.table(file.path(Tdrive, "data/curated/h2b.txt"),
                    stringsAsFactors = FALSE, sep = "\t") %>%
    rename("medname" = "V1") %>%
    mutate(medclass = "H2B")
  ppi <- read.table(file.path(Tdrive, "data/curated/proton-pump-inhibitors.txt"),
                    stringsAsFactors = FALSE, sep = "\t") %>%
    rename("medname" = "V1") %>%
    mutate(medclass = "PPI")
  bind_rows(h2b, ppi)
}

# Median offset between prescription ORDERING_DATE and START_DATE, used when
# start date is missing.
check_date_difference <- function(x) {
  tmp <- x %>%
    mutate(start = as.Date(START_DATE),
           ordering = as.Date(ORDERING_DATE)) %>%
    dplyr::select(start, ordering) %>%
    drop_na()
  summary(as.numeric(tmp$ordering - tmp$start))
}

class.key <- pull_class_assignments()
```

# Cohort 1 eligibility, survival time, and covariates

Age is calculated from date of birth and ICI start, then DOB is discarded.
OS time (`days`, `follow_up`) is days from ICI start to death or last contact;
calendar dates of death/contact are discarded.

```{r process-corr}
ppi_corr <- CoRR %>%
  rename("MRN" = "record_id") %>%
  filter(iostart != "") %>%
  filter(vitalstatus != 3) %>%
  mutate(vitalstatus = ifelse(vitalstatus == 1, 1, 0)) %>%
  mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
  mutate(date = coalesce(deathdate, lastcontact),
         date = as.Date(date),
         iostart = as.Date(iostart),
         dob = as.Date(dob)) %>%
  mutate(age = as.numeric(as.character(floor((iostart - dob) / 365)))) %>%
  mutate(follow_up = as.numeric(as.character(date - iostart))) %>%
  left_join(cancer.key) %>%
  mutate(
    Cancer = if_else(grepl("melanoma|non-small cell lung cancer", cancer.name),
                     true = cancer.name,
                     false = "Other"),
    Cancer = fct_relevel(Cancer, c("melanoma", "non-small cell lung cancer", "Other"))
  ) %>%
  filter(immunotherapy != 11) %>%
  drop_na(ppi, dob, iostart) %>%
  mutate(prior_chemo = ifelse(chemo_prior == 1, "yes", "no")) %>%
  mutate(eventdate = coalesce(deathdate, lastcontact)) %>%
  mutate(eventdate = as.POSIXct(eventdate, format = "%F"),
         iostart = as.POSIXct(iostart, format = "%F"),
         chemo_date = as.POSIXct(chemo_date, format = "%F")) %>%
  mutate(days = as.numeric(eventdate - iostart)) %>%
  mutate(chemo_timing = as.numeric(difftime(iostart, chemo_date, units = "days"))) %>%
  mutate(chemo_greater_one_year = ifelse(chemo_timing > 366, 1, 0))
```

# Medication timing relative to ICI start

For each PPI, H2B, and antibiotic prescription, compute the set of days
relative to ICI start that the patient was covered. These strings
(`PPI.days.to.iostart`, etc.) are used by the 30-day sliding-window analysis
(Figure 2E) and the ±28-day exposure flags (Figure 1C). Calendar prescription
dates are not retained.

```{r timing-functions}
pull_timing_info <- function(des.meds, meds.list, ord.offset, cdb, end.complete) {

  meds.members <- lapply(des.meds$medname, function(x) {
    as.data.frame(cbind(
      GENERIC_NAME = unique(meds.list$GENERIC_NAME[grepl(x, meds.list$GENERIC_NAME, ignore.case = TRUE)]),
      medname = x
    ))
  }) %>%
    bind_rows() %>%
    mutate(GENERIC_NAME = as.character(GENERIC_NAME),
           medname = as.character(medname))

  des.meds.full <- meds.members %>%
    left_join(des.meds) %>%
    filter(!is.na(GENERIC_NAME))

  iostarts <- cdb %>%
    select(MRN, iostart) %>%
    mutate(iostart = as.Date(iostart))

  time.obj.start <- des.meds.full %>%
    left_join(meds.list) %>%
    mutate(MRN = as.character(MRN),
           START_DATE = as.Date(START_DATE),
           ORDERING_DATE = as.Date(ORDERING_DATE),
           END_DATE = as.Date(END_DATE)) %>%
    select(MRN, GENERIC_NAME, START_DATE, ORDERING_DATE, END_DATE, REFILLS, medname, medclass) %>%
    inner_join(iostarts) %>%
    mutate(
      REFILLS = ifelse(is.na(REFILLS), 0, REFILLS),
      med.dur = ifelse(end.complete == TRUE & !is.na(END_DATE) & !is.na(START_DATE),
                       as.numeric(END_DATE - START_DATE + 1),
                       (REFILLS + 1) * 30),
      days.to.io = ifelse(is.na(START_DATE),
                          as.numeric(ORDERING_DATE - iostart) - ord.offset,
                          as.numeric(START_DATE - iostart)),
      stop.date = days.to.io + med.dur - 1
    )

  relative.days.covered <- lapply(
    1:nrow(time.obj.start),
    function(i) paste(time.obj.start$days.to.io[i]:time.obj.start$stop.date[i], collapse = ",")
  )
  time.obj.start$relative.days.covered <- unlist(relative.days.covered)

  time.class <- time.obj.start %>%
    group_by(MRN, medclass) %>%
    summarise(alldays = paste(relative.days.covered, collapse = ","), .groups = "drop")
  time.class.unq <- lapply(
    1:nrow(time.class),
    function(x) paste(unique(unlist(strsplit(time.class$alldays[x], split = ","))), collapse = ",")
  )
  time.class$alldays <- unlist(time.class.unq)
  time.class <- time.class %>%
    mutate(medclass = paste0(medclass, ".days.to.iostart")) %>%
    spread(key = medclass, value = alldays)

  cdb %>% left_join(time.class)
}

generate_abx_key <- function() {
  as.data.frame(cbind(
    medname = c(
      "penicillin", "Amoxicillin", "Ampicillin", "Piperacillin", "Cefalexin",
      "Cephalexin", "Ceftriaxone", "Ceftazidime", "Cefepime", "Aztreonam", "Imipenem",
      "Ciprofloxacin", "levofloxacin", "Moxifloxacin",
      "Azithromycin", "Clarithromycin", "Erythromycin",
      "Doxycycline", "Tetracycline", "Minocycline",
      "Linezolid", "Daptomycin",
      "Acyclovir", "Ganciclovir", "Valacyclovir", "Valaciclovir", "Valganciclovir",
      "Foscarnet", "Oseltamivir",
      "Fluconazole", "Voriconazole", "Posaconazole", "Caspofungin", "Micafungin",
      "Vancomycin", "Sulfamethoxazole", "Clindamycin", "Metronidazole"
    ),
    medclass = "ABX"
  ), stringsAsFactors = FALSE)
}

abx.key <- generate_abx_key()
start.ordering.offset <- check_date_difference(meds)

ppi_corr <- pull_timing_info(
  bind_rows(class.key, abx.key), meds,
  start.ordering.offset["Median"], ppi_corr, FALSE
)
```

# Collapse ICI regimens to the classes used in Table 1

```{r immunotherapy-labels}
ppi_corr <- ppi_corr %>%
  left_join(ionames) %>%
  mutate(
    Immunotherapy = if_else(grepl("Atezo|Durvalumab$|Nivolumab$|Pembro", Immunotherapy),
                            true = "PD1/L1", false = Immunotherapy),
    Immunotherapy = if_else(grepl("Durvalumab \\+ Tremelimumab$|Nivolumab \\+ Ipilimumab$", Immunotherapy),
                            true = "PD1/L1 + CTLA4", false = Immunotherapy),
    Immunotherapy = if_else(grepl("Tremelimumab$|Ipilimumab$", Immunotherapy),
                            true = "CTLA4", false = Immunotherapy),
    Immunotherapy = if_else(grepl("Nivolumab \\+ Chemotherapy", Immunotherapy),
                            true = "Other", false = Immunotherapy),
    Immunotherapy = fct_relevel(Immunotherapy, c("PD1/L1", "CTLA4", "PD1/L1 + CTLA4", "Other"))
  )
```

# Keep only variables used in the manuscript; drop PHI

Columns retained, and where they are used:

- Concomitant meds at ICI start (`ppi`, `statin`, `h2b_iostart`, `steroid_iostart`,
  `abx_iostart`, `nsaid_iostart`): Figure 1B, 1C, 2F, 2G, Figure S3
- Survival (`days`, `vitalstatus`): Figures 1D–F, 2A–G, Figure S3, Cox models
- Cancer labels (`cancer.name`, `Cancer`): Figure 1C heatmap; Table 1
- Relative medication-day strings (`PPI.days.to.iostart`, `H2B.days.to.iostart`,
  `ABX.days.to.iostart`): Figure 1C (±28-day window), Figure 2E (sliding window)
- Clinical covariates (`staging`, `cci_number`, `line_of_therapy`, `prior_chemo`,
  `chemo_greater_one_year`, `ecog`, `age`, `follow_up`, `sex`, `Immunotherapy`):
  Table 1 and stratified / multivariable survival models

Explicitly dropped: MRN, date of birth, death/last-contact dates, ICI start
calendar date, and all other REDCap fields not used in the paper.

```{r drop-phi}
ppi_corr <- ppi_corr %>%
  transmute(
    ppi, statin, h2b_iostart, steroid_iostart, abx_iostart, nsaid_iostart,
    vitalstatus, days,
    cancer.name, Cancer, Immunotherapy,
    PPI.days.to.iostart, H2B.days.to.iostart, ABX.days.to.iostart,
    staging, cci_number, line_of_therapy, prior_chemo, chemo_greater_one_year,
    ecog, age, follow_up, sex
  )
```

```{r write-cleaned}
out_file <- normalizePath(
  file.path(this_dir, "..", "data", "cohort1-cleaned-clinical.RDS"),
  mustWork = FALSE
)
dir.create(dirname(out_file), recursive = TRUE, showWarnings = FALSE)
saveRDS(ppi_corr, file = out_file)
```

```{r check-cleaned}
# Confirm the written object has no identifiers and only manuscript variables.
stopifnot(file.exists(out_file))
check <- readRDS(out_file)

forbidden <- c(
  "MRN", "record_id", "last_name", "dob", "iostart", "deathdate",
  "lastcontact", "eventdate", "chemo_date", "ppi_start", "date"
)
present_forbidden <- intersect(forbidden, names(check))
if (length(present_forbidden)) {
  stop("PHI-like columns still present: ", paste(present_forbidden, collapse = ", "))
}

message("Wrote: ", out_file)
message("Rows: ", nrow(check), "  Columns: ", ncol(check))
print(names(check))
```

