---
title: "Table 1 / Tables S2–S3 — cohort demographics"
author: "Anna Petouhoff"
date: "2025-02-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Directory of this notebook (for 00-paths.R). Table S3 still reads a
# de-identified spreadsheet from the lab share; Cohorts 1 and 2 do not.
this_dir <- if (!is.null(knitr::current_input())) {
  dirname(normalizePath(knitr::current_input()))
} else if (file.exists("Table 1.Rmd")) {
  getwd()
} else if (file.exists("manuscript/scripts/Table 1.Rmd")) {
  file.path(getwd(), "manuscript/scripts")
} else {
  getwd()
}

paths_file <- file.path(this_dir, "00-paths.R")
if (file.exists(paths_file)) {
  source(paths_file)
}
```

Produces the demographic tables:

- `../tables/table1_cohort1.csv` — **Table 1** (Cohort 1, all cancers)
- `../tables/table_cohort2.csv` — **Table S2** (Cohort 2, melanoma)
- `../tables/table_cohort3.csv` — **Table S3** (Cohort 3, NSCLC/RCC)

Only clinical variables that appear in those tables are read. Cohort 1 comes
from `../data/cohort1-cleaned-clinical.RDS` (written by `Processing.Rmd`).
Cohort 2 comes from `../data/Cohort_2/clinical.rds`. Cohort 3 Table S3 uses
the de-identified spreadsheet path `Iobio_clinical` from gitignored
`00-paths.R`.

```{r libraries}
library(dplyr)
library(tidyr)
library(tidyverse)
library(tableone)
library(forcats)
library(readxl)
```

```{r load}
# Table 1 variables only.
CoRR <- readRDS("../data/cohort1-cleaned-clinical.RDS") %>%
  select(ecog, staging, age, follow_up, sex, prior_chemo, ppi, vitalstatus, Cancer, Immunotherapy)

# Table S2 variables only (already trimmed in clinical.rds).
load("../data/Cohort_2/clinical.rds")
load("../data/Cohort_2/mitox_microbiome_metadata.rds")

# Table S3: de-identified Cohort 3 clinical file (lab share path from 00-paths.R).
if (!exists("Iobio_clinical") || !nzchar(Iobio_clinical) || !file.exists(Iobio_clinical)) {
  stop("Table S3 needs Iobio_clinical in 00-paths.R. See 00-paths.R.example.")
}
iobio <- read_excel(Iobio_clinical)
```

# Table 1 — Cohort 1

```{r table1-cohort1}
CoRR_df <- CoRR %>%
  mutate(
    `ECOG PS` = fct_relevel(ifelse(ecog > 2, ">2", ecog), "0", "1", "2", ">2"),
    Staging = ifelse(staging == 5, "Unknown", staging),
    Age = age,
    follow_up_time = follow_up,
    Sex = sex,
    Chemo_prior = prior_chemo,
    Immunotherapy = fct_relevel(Immunotherapy, "CTLA4", "PD1/L1", "PD1/L1 + CTLA4", "Other")
  )

listVars <- c("Age", "Sex", "vitalstatus", "ECOG PS", "Cancer", "Staging",
              "Immunotherapy", "Chemo_prior", "follow_up_time")
catVars <- c("Sex", "Cancer", "vitalstatus", "Immunotherapy", "ECOG PS",
             "Chemo_prior", "Staging")
table1 <- CreateTableOne(vars = listVars, data = CoRR_df, factorVars = catVars,
                         strata = c("ppi"))
tbl1.ppi <- print(table1, nonnormal = "follow_up_time")
write.csv(tbl1.ppi, "../tables/table1_cohort1.csv")
```

# Table S2 — Cohort 2 (baseline microbiome samples within ±10 days of ICI start)

```{r table-s2}
baseline <- mic.meta %>%
  group_by(`Patient Id`) %>%
  arrange(Days) %>%
  slice_head(n = 1) %>%
  ungroup() %>%
  filter(abs(Days) <= 10)

demo.filt <- demographics %>%
  select(`Patient Id`, Age, Sex, Treatment, Stage, PPI) %>%
  filter(`Patient Id` %in% baseline$`Patient Id`) %>%
  mutate(PPI = if_else(is.na(PPI), "No", as.character(PPI)))

listVars3 <- c("Age", "Sex", "Treatment", "Stage")
table3 <- CreateTableOne(vars = listVars3, data = demo.filt, strata = c("PPI"))
write.csv(print(table3), "../tables/table_cohort2.csv")
```

# Table S3 — Cohort 3

```{r table-s3}
iobio <- iobio %>%
  rename(
    Immunotherapy = "Type of immunotherapy",
    ECOG = "ECOG at start date",
    VitalStatus = "What is the patient's vital status? 1-deceaased",
    PPI = "PPI Y/N"
  ) %>%
  select(Gender, VitalStatus, ECOG, Immunotherapy, PPI)

iobio$Immunotherapy <- as.factor(iobio$Immunotherapy)

listVars4 <- c("Gender", "VitalStatus", "ECOG", "Immunotherapy")
catVars4 <- c("Gender", "VitalStatus", "Immunotherapy", "ECOG")
table4 <- CreateTableOne(vars = listVars4, data = iobio, factorVars = catVars4,
                         strata = c("PPI"))
write.csv(print(table4), "../tables/table_cohort3.csv")
```
