---
title: "CoRR 1st Line LC"
author: "Anna Petouhoff"
date: "2025-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(survival)
library(survminer)
library(svglite)
library(patchwork)
```

```{r}
CoRR <- read.csv("T:/Labs/Spakowicz/ppiio/data/derived/2025-02-17/CoRR_formatted.csv")
P188 <- read.csv("T:/Labs/Spakowicz/ppiio/data/derived/2025-02-17/P188_formatted.csv")
p188_34 <- read.csv("T:/Labs/Spakowicz/ppiio/data/derived/2025-02-17/P188_formatted_Stage_34.csv") 
```

```{r}
CoRR <- CoRR %>%
  filter(line_of_therapy != 1) %>%
  filter(cancer == 1) %>%
  drop_na(prior_chemo)
```


# PPI Use and OS across all cancers 
```{r}
all_surv <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR)

all_surv_plot <- ggsurvplot(all_surv, pval = T, risk.table = T) +
  ggtitle("PPI use and OS across all cancers")

combined_all_surv <- all_surv_plot$plot / all_surv_plot$table + plot_layout(heights = c(3, 1))
combined_all_surv
```

# PPI use and OS without prior chemotherapy
```{r}
no_chemo <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$prior_chemo == "no",])

no_chemo_plot <- ggsurvplot(no_chemo, pval = T, risk.table = T) +
  ggtitle("PPI use and OS without prior chemotherapy")

combined_no_chemo <- no_chemo_plot$plot / no_chemo_plot$table + plot_layout(heights = c(3, 1))
combined_no_chemo
```
# PPI use and OS with prior chemotherapy
```{r}
chemo <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$prior_chemo == "yes",])

chemo_plot <- ggsurvplot(chemo, pval = T, risk.table = T) +
  ggtitle("PPI use and OS with prior chemotherapy")

combined_chemo <- chemo_plot$plot / chemo_plot$table + plot_layout(heights = c(3, 1))
combined_chemo
```

# PPI use and OS if patient recieved chemo more than 1 year

```{r}
prior_chemotherapy <- CoRR %>%
  drop_na(chemo_greater_one_year)

more_than_one_year <- survfit(Surv(days, vitalstatus) ~ ppi, data = prior_chemotherapy[prior_chemotherapy$chemo_greater_one_year == 1,])

more_than_one_year_plot <- ggsurvplot(more_than_one_year, pval = T, risk.table = T) +
  ggtitle("PPI use and OS if patient recieved chemo more than 1 year")

combined_more_than_one_year <- more_than_one_year_plot$plot / more_than_one_year_plot$table + plot_layout(heights = c(3, 1))
combined_more_than_one_year
```

# PPI use and OS if patient recieved chemo less than 1 year
```{r}
less_than_one_year <- survfit(Surv(days, vitalstatus) ~ ppi, data = prior_chemotherapy[prior_chemotherapy$chemo_greater_one_year == 0,])

less_than_one_year_plot <- ggsurvplot(less_than_one_year, pval = T, risk.table = T) +
  ggtitle("PPI use and OS if patient recieved chemo less than 1 year")

combined_less_than_one_year <- less_than_one_year_plot$plot / less_than_one_year_plot$table + plot_layout(heights = c(3, 1))
combined_less_than_one_year
```
```{r}
PDL1 <- survfit(Surv(days, vitalstatus) ~ ppi, data = CoRR[CoRR$Immunotherapy == "PD1/L1",])

PDL1_plot <- ggsurvplot(PDL1, pval = T, risk.table = T) +
  ggtitle("P188: PPI use and OS for PD1/L1")

combined_PDL1 <- PDL1_plot$plot / PDL1_plot$table + plot_layout(heights = c(3, 1))
combined_PDL1
```

```{r}
filtered <- CoRR %>%
  filter(Immunotherapy == "PD1/L1", chemo_greater_one_year == 1)
PDL1_Chemo <- survfit(Surv(days, vitalstatus) ~ ppi, data = filtered)
PDL1_Chemo_plot <- ggsurvplot(PDL1_Chemo, pval = T, risk.table = T) +
  ggtitle("P188: PPI use and OS for PD1/L1 with <1 year chemo")

combined_PDL1_Chemo <- PDL1_Chemo_plot$plot / PDL1_Chemo_plot$table + plot_layout(heights = c(3, 1))
combined_PDL1_Chemo
```



ggsave("../../figures/2025/PPI_all_cancers_KM.svg", plot = combined_all_surv, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_first_line_KM.svg", plot = combined_first_line, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_second_line_KM.svg", plot = combined_second_line, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_no_chemo_KM.svg", plot = combined_no_chemo, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_chemo_KM.svg", plot = combined_chemo, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_more_than_one_year_KM.svg", plot = combined_more_than_one_year, device = "svg", width = 6, height = 6)
ggsave("../../figures/2025/PPI_less_than_one_year.svg", plot = combined_less_than_one_year, device = "svg", width = 6, height = 6)

```{r}
P188 <- P188 %>%
  mutate(filter_date = substr(iostopdate, 1, 4)) %>%
  filter(!(filter_date %in% c("2021", "2022")))
```

