---
title: "Processing"
author: "Anna Petouhoff"
date: "2025-02-17"
output: html_document
---

```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(readxl)
```

# read in csv
```{r}
CoRR <- read.csv("T:/Labs/Spakowicz/ppiio/data/raw/CoRR4684/2025-02-17/CoRR4684ARetrospecti_DATA_2025-02-17_2009.csv")
P188 <- read.csv("T:/Labs/Spakowicz/ppiio/data/raw/P188/2025-02-17/P188NSCLCIOTox111611_DATA_2025-02-17_1010.csv")
```

#create a cancer key 
```{r}
cancer.copy <- c("1 non-small cell lung cancer
2 small cell lung cancer
3 melanoma
4 renal cell carcinoma
5 head and neck carcinoma
6 merkel cell carcinoma
7 Hodgkin lymphoma
8 breast cancer
9 colon cancer
10 pancreatic cancer
11 Sarcoma
12 Prostate
13 bladder cancer
14 AML
15 ALL
17 Other") %>%
  strsplit(., split = "\n") %>%
  unlist

cancer.key <- 
  data.frame(
    cancer =  cancer.copy %>%
      gsub("(\\d+).*", "\\1", .) %>%
      as.integer(),
    cancer.name = cancer.copy %>%
      gsub("\\d+ (.*)", "\\1", .),
    stringsAsFactors = FALSE
  )
```


# process CoRR4684
```{r}
ppi_corr <- CoRR %>%
   filter(iostart != "") %>%
  mutate(vitalstatus = ifelse(vitalstatus == 1, 1, 0)) %>%
  mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
  mutate(date = coalesce(deathdate, lastcontact),
         date = as.Date(date),
         iostart = as.Date(iostart),
         dob = as.Date(dob)) %>%
  mutate(age=as.numeric(as.character(floor((date-dob)/365))))%>%
   mutate(follow_up = as.numeric(as.character(date- iostart)))%>%
  left_join(cancer.key) %>%
  mutate(Cancer = if_else(grepl("melanoma|non-small cell lung cancer",
                                cancer.name),
                          true = cancer.name,
                          false = "Other"),
         Cancer = fct_relevel(Cancer, c("melanoma", "non-small cell lung cancer",
                                        "Other"))) %>%
  filter(immunotherapy != 11) %>%
  drop_na(ppi)

```

```{r}
write.csv(ppi_corr,  "T:/Labs/Spakowicz/ppiio/data/derived/2025-02-17/CoRR_formatted.csv")
```


# Process P188
```{r}
ppi_P188 <- P188 %>%
  filter(cancer == 1) %>%
  filter(staging == 4) %>%
  filter(iostart != "")  %>%
  mutate(vitalstatus = ifelse(vitalstatus == 1, 1, 0)) %>%
  mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
  mutate(date = coalesce(deathdate, lastcontact),
         date = as.Date(date),
         iostart = as.Date(iostart),
         dob = as.Date(dob)) %>%
  mutate(age=as.numeric(as.character(floor((date-dob)/365))))%>%
   mutate(follow_up = as.numeric(as.character(date- iostart)))%>%
 drop_na(ppi)
```

```{r}
write.csv(ppi_P188, "T:/Labs/Spakowicz/ppiio/data/derived/2025-02-17/P188_formatted.csv")
```


ppi_corr <- CoRR %>%
  rename("MRN"="record_id") %>%
  filter(vitalstatus != 3) %>%
  mutate(vitalstatus = ifelse(vitalstatus == 1, 1, 0))%>%
  mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
  mutate(date = coalesce(deathdate, lastcontact),
         date = as.Date(date),
         iostart = as.Date(iostart),
         dob = as.Date(dob))%>%
  mutate(age=as.numeric(as.character(floor((date-dob)/365))))%>%
  filter(!is.na(iostart))%>%
  mutate(follow_up = as.numeric(as.character(date- iostart)))%>%
  left_join(cancer.key) %>%
  mutate(Cancer = if_else(grepl("melanoma|non-small cell lung cancer",
                                cancer.name),
                          true = cancer.name,
                          false = "Other"),
         Cancer = fct_relevel(Cancer, c("melanoma", "non-small cell lung cancer",
                                        "Other"))) %>%
  filter(immunotherapy != 11) %>%
  drop_na(iostart, ppi, date, dob) %>%
  filter(!MRN %in% noio)
