---
title: "Figure-2_details-of-significant-meds"
author: "Rebecca Hoyd"
date: "May 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(survminer)
library(glue)
library(RColorBrewer)
library(forcats)
library(broom)
library(ggdendro)
library(gridExtra)
```

```{r load data}
# Select source
source("00-paths.R")
# Load in data frame
db <- readRDS(file.path(paths$ppiio_derived, "deidentdb_timing.RDS"))

# May not need this
sterkey <- read.table("C:/Users/muni11/Box/abio/manuscript/data/from_processing/steroid-dbid-key.tsv") %>%
  select(-Drug.name)

sterkey <- sterkey[!duplicated(sterkey),]

```

```{r}
# abxclasses <- c("BetaLactam", "Macrolide",  "Other.Antibacterial", "Fluoroquinolone", "Tetracycline",
#                 "Vancomycin", "Sulfamethoxazole", "Clindamycin", "Metronidazole")
# 
# orgtype <- c("Antibacterial", "Antiviral", "Antifungal")

allcancers <- c('Other', 'Melanoma', 'Non-Small Cell Lung Cancer', 'Head and Neck Carcinoma', 'Renal Cell Carcinoma', 'Sarcoma', 'Bladder Cancer')
```

# A) Hazard Ratio over time

```{r}
check.window <- function(vect, per){
  wind <- per[1]:per[2]
  wind <- glue_collapse(as.character(wind), sep = ",|,")
  wind <- paste(",", wind, ",", sep = "")
  ifelse(grepl(pattern = wind, x = vect), 1, 0)
}
```

```{r}
db <- db %>%
  # Add commas to work with 'check.window'
  mutate(ppidays.to.iostart = paste(",", ppidays.to.iostart, ",", sep = ""),
         # Run 'check.window' to create a binary column telling us if ppi use fell within 90 days of IO
         ppi.90.90 = check.window(ppidays.to.iostart, c(-90,90)))
```


# I am not sure what this is coding for an I cannot get it to run correctly--even with the original code
## Part 1: Steroids






*Error in `$<-.data.frame`(`*tmp*`, tmp2, value = logical(0)) : replacement has 0 rows, data has 119
```{r, include = FALSE}
ppifloattime <- db %>%
  filter(!is.na(ppi_start) & ppi_start != "") %>%
  select(days, vitalstatus, ppidays.to.iostart)
  

ppi.zero.cohort <- ppifloattime %>%
  mutate(tmp = check.window(ppidays.to.iostart, c(-180,180))) %>%
  filter(tmp == 0) %>%
  select(days, vitalstatus, tmp)

leftanchor <-c(-180:151)
ppiwindresults <- list()

i <- 1
ppifloattime$tmp2 <- check.window(ppifloattime$ppidays.to.iostart, c(leftanchor[i], leftanchor[i] + 29))

for(i in 1:length(leftanchor)){
  j <- leftanchor[i]
  ppifloattime$tmp <- check.window(ppifloattime$ppidays.to.iostart, c(j, j+29))
  
  ppifloattime <- ppifloattime %>%
    mutate(tmp3 = tmp2 - tmp) %>%
    filter(tmp3 == 0 | tmp3 == -1) %>%
    mutate(tmp2 = tmp)
  
  ppifloatforsurv <- ppifloattime %>%
    filter(tmp == 1) %>%
    bind_rows(ppi.zero.cohort)
  
  try({
    survtemp <- coxph(Surv(days, vitalstatus) ~ tmp, data = ppifloatforsurv)
    ppiwindresults[[i]] <- survtemp %>%
      tidy() %>%
      mutate(hazard.ratio = exp(estimate),
             # low.bound = exp(conf.low),
             # upper.bound = exp(conf.high),
             windpval = p.value,
             leftanchor = j) %>%
      select(hazard.ratio, 
             # low.bound, upper.bound, 
             leftanchor, windpval, std.error)
  })
  
  
}

ppipvaltime <- bind_rows(ppiwindresults) %>%
   mutate(leftanchor = leftanchor + 15)
```

## Part 2: Antibiotics

```{r, message=FALSE, error=FALSE, warning=FALSE}
abxfloattime <- db %>%
  select(days, vitalstatus, abxdays.to.iostart) %>%
  mutate(abx.days.to.iostart = paste(",", abxdays.to.iostart, ",", sep = ""))

abx.zero.cohort <- db %>%
  select(days, vitalstatus, abx.180.180) %>%
  filter(abx.180.180 == 0) %>%
  mutate(tmp = abx.180.180) %>%
  select(days, vitalstatus, tmp)

leftanchor <-c(-180:151)
abxwindresults <- list()

i <- 1
abxfloattime$tmp2 <- check.window(abxfloattime$abxdays.to.iostart, c(leftanchor[i], leftanchor[i]+29))
  
for(i in 1:length(leftanchor)){
  j <- leftanchor[i]
  abxfloattime$tmp <- check.window(abxfloattime$abxdays.to.iostart, c(j, j+29))
  
  abxfloattime <- abxfloattime %>%
    mutate(tmp3 = tmp2 -tmp) %>%
    filter(tmp3 == 0 | tmp3 == -1) %>%
    mutate(tmp2 = tmp)
  
  abxfloatforsurv <- abxfloattime %>%
    filter(tmp == 1) %>%
    bind_rows(abx.zero.cohort)
  
  try({
  survtemp <- coxph(Surv(days, vitalstatus) ~ tmp, data = abxfloatforsurv)
  abxwindresults[[i]] <- survtemp %>%
    tidy() %>%
    mutate(hazard.ratio = exp(estimate),
           low.bound = exp(conf.low),
           upper.bound = exp(conf.high), 
           windpval = p.value,
           leftanchor = j) %>%
    select(hazard.ratio, low.bound, upper.bound, windpval, leftanchor)
  })
  
}

abxpvaltime <- bind_rows(abxwindresults)
abxpvaltime$leftanchor <- abxpvaltime$leftanchor + 15
```


## Combine + plot results

```{r}
abxpvaltime$source <- "ABx"
sterpvaltime$source <- "CS"
allpvaltime <- bind_rows(abxpvaltime, sterpvaltime)

ggplot(data = allpvaltime, aes(x = leftanchor, y = hazard.ratio)) +
  geom_line(aes(color = source)) +
  geom_ribbon(aes(ymin = low.bound, ymax = upper.bound, fill = source), alpha = 0.5) +
  geom_hline(yintercept = 1) +
  scale_color_manual(aesthetics = c("fill", "color"), breaks = c("ABx", "CS"), 
                     values = c("darkred", "darkblue"), name = "Drug") +
  labs(title = "", x = "Days relative to ICI start",y = "Hazard Ratio" ) +
  theme_bw() + 
  theme(axis.text = element_text(size = 12), axis.title = element_text(size = 15)) 
```


# B) P.value heatmaps: strongest window

## Create window columns:
```{r}
statesofinterest <- c(abxclasses, orgtype, as.character(sterkey$Generic.ID))

edges <- c(-28,28)

db.heat <- db %>%
  select(days, cancer.name, vitalstatus, one_of(paste(statesofinterest, ".to.start", sep = "")))

for(s in statesofinterest){
  db.heat[[paste(s, ".to.start", sep = "")]] <- paste(",", db.heat[[paste(s, ".to.start", sep = "")]], ",", 
                                                      sep = "") 
  db.heat[[s]] <- check.window(db.heat[[paste(s, ".to.start", sep = "")]], edges)
}

db.heat$Other.Steroid <- ifelse(db.heat$Betamthasone == 1|db.heat$Fludrocortisone == 1|db.heat$Prednisolone == 1,
                                1, 0)
```

## Prepare data for the antibiotic heatmaps

Loop to run all coxph() models and to save other data for the heatmap figures:
```{r}
statesofinterest <- c(abxclasses, orgtype)

curvesbycancer <- list()
sfits <- list()

pvalbycancbac <- list()
pvalbycancorg <- list()

hrbycancbac <- list()
hrbycancorg <- list()

labcolbac <- list()
labcolorg <- list()


g <- as.formula(paste("Surv(days, vitalstatus) ~ ", glue_collapse(abxclasses, sep = " + "), sep = ""))
h <- as.formula(paste("Surv(days, vitalstatus) ~ ", glue_collapse(orgtype, sep = " + "), sep = ""))

for(c in allcancers){
  
  tmp <- subset(db.heat, db.heat$cancer.name == c)
  
  x <- coefficients(summary(coxph(formula = g, data = tmp)))
  y <- coefficients(summary(coxph(formula = h, data = tmp)))

    hrbycancbac[[c]] <- round(x[,2], 2)
    labcolbac[[c]] <- ifelse(x[,2] < 1, "red", "black")
    pvalbycancbac[[c]] <- x[,5]
    
    hrbycancorg[[c]] <- round(y[,2], 2)
    labcolorg[[c]] <- ifelse(y[,2] < 1, "red", "black")
    pvalbycancorg[[c]] <- y[,5]
  
}

perorgcancpvals <- bind_rows(pvalbycancorg) %>%
  t() %>%
  as.data.frame()
colnames(perorgcancpvals) <- names(pvalbycancorg$Other)

perclasscancpvals <- bind_rows(pvalbycancbac) %>%
  t() %>%
  as.data.frame()
colnames(perclasscancpvals) <- names(pvalbycancbac$Other)


labcolbac <- bind_rows(labcolbac) %>%
  t() %>%
  as.data.frame()
colnames(labcolbac) <- names(hrbycancbac$Other)

labcolorg <- bind_rows(labcolorg) %>%
  t() %>%
  as.data.frame()
colnames(labcolorg) <- names(hrbycancorg$Other)

```

Code manipulating data into long format for the creation of figures, and saving a data frame object that will be used for the heatmap.
```{r}

labcolorg <- labcolorg %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "AbxClass", value = "labcol", -cancer) %>%
  mutate(labcol = ifelse(is.na(labcol), "black", labcol))

labcolbac <- labcolbac %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "AbxClass", value = "labcol", -cancer) %>%
  mutate(labcol = ifelse(is.na(labcol), "black", labcol))

heatorg <- as.data.frame(perorgcancpvals)  %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "AbxClass", value = "p.value", -cancer) %>%
  full_join(labcolorg)

heatclass <- as.data.frame(perclasscancpvals)  %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "AbxClass", value = "p.value", -cancer) %>%
  full_join(labcolbac)

heatclass <- bind_rows(heatorg, heatclass)
```

Save data from all cancers:
```{r}

x <- coefficients(summary(coxph(formula = g, data = db.heat)))
y <- coefficients(summary(coxph(formula = h, data = db.heat)))
albac.hr <- c(round(x[,2], 2), round(y[,2], 2))
albac.pval<- c(x[,5], y[,5])


albac.dat <- as.data.frame(cbind(albac.hr, albac.pval))
colnames(albac.dat) <- c("hazrat", "p.value")

albac.dat <- albac.dat %>%
  rownames_to_column(var = "AbxClass") %>%
  mutate(labcol = ifelse(hazrat < 1, "red", "black"),
         cancer = "All") %>%
  select(cancer, AbxClass, p.value, labcol)
```


Save data from all meds:
```{r}
sfits <- list()
db <- db %>%
  mutate(abx.28.28 = check.window(paste(",", abxdays.to.iostart, ",", sep = ""), c(-28,28)))

for(c in allcancers){
  tmp <- db %>%
    filter(cancer.name == c)
  
  sfits[[c]] <- surv_pvalue(survfit(Surv(days, vitalstatus) ~ abx.28.28, data = tmp))[1,2]
}

alcancbac.pval <- bind_rows(sfits)
alcancbac.hr <- lapply(allcancers, function(x) { 
  coefficients(summary(coxph(Surv(days, vitalstatus) ~ abx.28.28,
                             data = filter(db, cancer.name == x))))[,2] 
  })

names(alcancbac.hr) <- allcancers

alcancbac.pval <- bind_rows(alcancbac.pval)
alcancbac.hr <- bind_rows(alcancbac.hr)

alcancbac.dat <- as.data.frame(bind_rows(alcancbac.pval, alcancbac.hr))
rownames(alcancbac.dat) <- c("p.value", "hazrat")

alcancbac.dat <- alcancbac.dat %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "cancer") %>%
  mutate(labcol = ifelse(hazrat < 1, "red", "black"),
         AbxClass = "All") %>%
  select(cancer, AbxClass, p.value, labcol)
  
```

the all/all corner:
```{r}
sfit <- survfit(Surv(days, vitalstatus) ~ abx.28.28, data = db)

allallbac <- as.data.frame(cbind(AbxClass = "All", cancer = "All", 
                                 p.value = surv_pvalue(sfit)[1,2], 
                                 hazrat = coefficients(summary(coxph(Surv(days, vitalstatus) ~ abx.28.28,
                                                                     data = db)))[1,2])) %>%
  mutate(p.value = as.numeric(as.character(p.value)),
         hazrat = as.numeric(as.character(hazrat)),
         labcol = ifelse(hazrat < 1, "red", "black")) %>%
  select(cancer, AbxClass, p.value, labcol)
```

Adding preserved full cohort:
```{r}
heatclass <- bind_rows(heatclass, albac.dat, alcancbac.dat, allallbac)
```

## Prepare data for the steroid heatmaps

```{r, message=FALSE, warning=FALSE}
sterdbid <- c("Triamcinolone", "Prednisone", "Hydrocortisone", "Methylprednisone", "Dexamethosone", "Other.Steroid")

dbidcurvesbycancer <- list()

dbidpvalbycancer <- list()

dbidhrbycanc <- list()

dbidlabcol <- list()

g <- as.formula(paste("Surv(days, vitalstatus) ~ ", glue_collapse(sterdbid, sep = " + "), sep = ""))

for (c in allcancers) {
  tmp <- subset(db.heat, db.heat$cancer.name == c)
  
  x <- coefficients(summary(coxph(formula = g, data = tmp)))
  dbidhrbycanc[[c]] <- round(x[,2], 2)
  dbidlabcol[[c]] <- ifelse(x[,2] < 1, "red", "black")
  dbidpvalbycancer[[c]] <- x[,5]
  
}


dbidallcancpvals <- bind_rows(dbidpvalbycancer) %>%
  t() %>%
  as.data.frame()
colnames(dbidallcancpvals) <- names(dbidpvalbycancer$Other)


dbidlabcol <- bind_rows(dbidlabcol) %>%
  t() %>%
  as.data.frame()
colnames(dbidlabcol) <- names(dbidhrbycanc$Other)

```


Manipulatin
```{r}

dbidlabcol <- dbidlabcol %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "Steroid.ID", value = "labcol", -cancer) %>%
  mutate(labcol = ifelse(is.na(labcol), "black", labcol))

heatster <- as.data.frame(dbidallcancpvals)  %>%
  rownames_to_column(var = "cancer") %>%
  gather(key = "Steroid.ID", value = "p.value", -cancer) %>%
  full_join(dbidlabcol) 
```

Save data from all cancers:
```{r}
forms <- list()
for(s in sterdbid){
  forms[[s]] <- as.formula(paste("Surv(days, vitalstatus) ~ ", s, sep = ""))
}
sfits <- surv_fit(forms, data = db.heat)
alster.pval <- lapply(sfits, function (x) surv_pvalue(x)[1,2])
names(alster.pval) <- names(forms)
alster.hr <- lapply(forms, function(x)  coefficients(summary(coxph(formula = x, data = db.heat)))[,2])

alster.pval <- bind_rows(alster.pval)
alster.hr <- bind_rows(alster.hr)

alster.dat <- as.data.frame(bind_rows(alster.pval, alster.hr))
rownames(alster.dat) <- c("p.value", "hazrat")

alster.dat <- alster.dat %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Steroid.ID") %>%
  mutate(labcol = ifelse(hazrat < 1, "red", "black"),
         cancer = "All") %>%
  select(cancer, Steroid.ID, p.value, labcol)
  
```


Save data from all meds:
```{r}
sfits <- list()
db <- db %>%
  mutate(ster.28.28 = check.window(paste(",", sterdays.to.start, ",", sep = ""), c(-28,28)))

for(c in allcancers){
  tmp <- db %>%
    filter(cancer.name == c)
  
  sfits[[c]] <- survfit(Surv(days, vitalstatus) ~ ster.28.28, data = tmp)
}

alcancster.pval <- lapply(sfits, function (x) surv_pvalue(x)[1,2])
alcancster.hr <- lapply(allcancers, function(x) { 
  coefficients(summary(coxph(Surv(days, vitalstatus) ~ ster.28.28,
                             data = filter(db, cancer.name == x))))[,2]
  })

names(alcancster.hr) <- allcancers

alcancster.pval <- bind_rows(alcancster.pval)
alcancster.hr <- bind_rows(alcancster.hr)

alcancster.dat <- as.data.frame(bind_rows(alcancster.pval, alcancster.hr))
rownames(alcancster.dat) <- c("p.value", "hazrat")

alcancster.dat <- alcancster.dat %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "cancer") %>%
  mutate(labcol = ifelse(hazrat < 1, "red", "black"),
         Steroid.ID = "All") %>%
  select(cancer, Steroid.ID, p.value, labcol)
  
```

the all/all corner:
```{r}
sfit <- survfit(Surv(days, vitalstatus) ~ ster.28.28, data = db)

allallster <- as.data.frame(cbind(Steroid.ID = "All", cancer = "All",
                                  p.value = surv_pvalue(sfit)[1,2], 
                                  hazrat = coefficients(summary(coxph(Surv(days, vitalstatus) ~ ster.28.28,
                                                                      data = db)))[1,2])) %>%
  mutate(p.value = as.numeric(as.character(p.value)),
         hazrat = as.numeric(as.character(hazrat)),
         labcol = ifelse(hazrat < 1, "red", "black")) %>%
  select(cancer, Steroid.ID, p.value, labcol)
```

Adding preserved full cohort:
```{r}
heatster <- bind_rows(heatster, alster.dat, alcancster.dat, allallster)
```

## Create dendrograms for the heatmaps

Antibacerials:
```{r}
hrforclustbac <- bind_rows(hrbycancbac) %>%
  as.data.frame()

rownames(hrforclustbac) <- names(hrbycancbac$Other)

hrforclustbac <- sapply(hrforclustbac, function(x) ifelse(x > 65000, 65000, x))
hrforclustbac <- apply(hrforclustbac, 2, function(x) ifelse(is.na(x), 1, x))
rownames(hrforclustbac) <- names(hrbycancbac$Other)
hrforclustbac <- dist(hrforclustbac)

bactree <- hclust(hrforclustbac)
bactree <- as.dendrogram(bactree)
bactree <- dendro_data(bactree)

bacorder <- as.character(bactree$labels$label)


dendplot <- ggdendrogram(bactree, rotate = FALSE, leaf_labels = FALSE) +
  geom_segment(aes(x = 1, xend = 1, y = -4000, yend = 0)) +
  geom_segment(aes(x = 2, xend = 2, y = -4000, yend = 0)) +
  geom_segment(aes(x = 3, xend = 3, y = -4000, yend = 0)) +
  geom_segment(aes(x = 4, xend = 4, y = -4000, yend = 0)) +
  geom_segment(aes(x = 5, xend = 5, y = -4000, yend = 0)) +
  geom_segment(aes(x = 6, xend = 6, y = -4000, yend = 0)) +
  geom_segment(aes(x = 7, xend = 7, y = -4000, yend = 0)) +
  geom_segment(aes(x = 8, xend = 8, y = -4000, yend = 0)) +
  geom_segment(aes(x = 9, xend = 9, y = -4000, yend = 0)) +
  theme(axis.text.y = element_blank(), axis.text.x = element_blank(),plot.margin = margin(0,0,0,0))

cancclust <- bind_rows(hrbycancbac) %>%
  t() %>%
  as.data.frame

cancclust <- sapply(cancclust, function(x) ifelse(x > 65000, 65000, x))
cancclust <- apply(cancclust, 2, function(x) ifelse(is.na(x), 1, x))
colnames(cancclust) <- names(hrbycancbac$Other)
rownames(cancclust) <- names(hrbycancbac)

cancclust <- dist(cancclust)
cancclust <- hclust(cancclust)
canctree <- dendro_data(as.dendrogram(cancclust))
```

Steroids:
```{r}

dbidforclust <- bind_rows(dbidhrbycanc) %>%
  as.data.frame() 

rownames(dbidforclust) <- names(dbidhrbycanc$Other)

dbidforclust <- sapply(dbidforclust, function(x) ifelse(x > 65000, 65000, x))
dbidforclust <- apply(dbidforclust, 2, function(x) ifelse(is.na(x), 1, x))

rownames(dbidforclust) <- names(dbidhrbycanc$Other)

dbidsterclust <- dist(dbidforclust)
dbidtree <- hclust(dbidsterclust)
dbidtree <- dendro_data(as.dendrogram(dbidtree))

dbiddendplot <- ggdendrogram(dbidtree, rotate = FALSE, size = 10, labels = FALSE) +
  theme(axis.text.y = element_blank(), axis.text.x = element_blank())

dbidcancclust <- t(dbidforclust) %>%
  dist() %>%
  hclust()
dbidcanctree <- dendro_data(as.dendrogram(dbidcancclust))
```

# Construct Heatmaps

Build the heatmap for the antibacterials:
```{r}
heatplot <- heatclass %>%
  ggplot(aes(fct_relevel(AbxClass, "All", "Antibacterial", "Antifungal", "Antiviral", bacorder), 
             fct_relevel(cancer, "Other", "Bladder Cancer", "Head and Neck Carcinoma", "Melanoma",
                         "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma", "Sarcoma", "All"))) +
  geom_tile(aes(fill = p.value )) +
  geom_point(aes(shape = labcol), size = 5) +
  scale_shape_manual(breaks = c("black", "red"), values = c("-", "+"), 
                     labels = c("HR>1", "HR<1"), name = "Direction") +
  labs(x = "", y = "") +
  scale_fill_distiller(guide = "legend", breaks = c(0.000000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1),
                       limits = c(0, 1), values = c(0.000000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1),
                       type = "div", palette = "RdBu", direction = 1, 
                       labels = c("<0.01", "0.01", "0.02","0.03", "0.04","0.05","0.1", "0.2", "0.5", "1")) +
  scale_x_discrete(breaks = c("All", "Antifungal", "Antiviral", "Antibacterial", "BetaLactam", "Macrolide",
                              "Other.Antibacterial", "Fluoroquinolone", "Tetracycline", "Vancomycin",
                              "Sulfamethoxazole", "Clindamycin", "Metronidazole"),
                   labels =c("All", "Antifungal", "Antiviral", "Antibacterial", "BetaLactam", "Macrolide", 
                             "Linezolid", "Fluoroquinolone", "Tetracycline", "Vancomycin", "Sulfamethoxazole-TMP",
                             "Clindamycin", "Metronidazole")) +
  scale_y_discrete(breaks = c("Other", "Bladder Cancer", "Head and Neck Carcinoma", "Melanoma",
                              "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma", "Sarcoma", "All"),
                   labels = c("Other", "Bladder Cancer", "Head and Neck\nCarcinoma", "Melanoma", 
                              "Non-Small Cell\nLung Cancer", "Renal Cell\nCarcinoma", "Sarcoma", "All")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12),
        axis.text = element_text(size = 12), legend.position = "none", plot.margin = margin(0,0,0,0))


```

Construct heatmap for steroids:
```{r}
dbidheatplot <- heatster %>%
  ggplot(aes(fct_relevel(Steroid.ID, c("All", as.character(dbidtree$labels$label))), 
             fct_relevel(cancer, "Other", "Bladder Cancer", "Head and Neck Carcinoma", "Melanoma", 
                         "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma", "Sarcoma", "All"))) +
  geom_tile(aes(fill = p.value)) +
  geom_point(aes(shape = labcol), size = 5) +
  scale_shape_manual(breaks = c("black", "red"), values = c("-", "+"), 
                     labels = c("HR>1", "HR<1"), name = "Direction") +
  labs(x = "", y = "") +
  scale_fill_distiller(guide = "legend", limits = c(0, 1),
                       breaks = c(0.00000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1),
                       values = c(0.00000000001, 0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2, 0.5, 1), 
                       labels = c("<0.01", "0.01", "0.02","0.03", "0.04","0.05","0.1", "0.2", "0.5", "1"),
                       type = "div", palette = "RdBu", direction = 1) +
  scale_x_discrete(breaks = c("All", "Methylprednisone", "Prednisone", "Dexamethosone", "Hydrocortisone",
                              "Other.Steroid", "Triamcinolone"), 
                   labels = c("All", "Methylprednisone", "Prednisone", "Dexamethosone", "Hydrocortisone",
                              "Other Steroid", "Triamcinolone")) +
  scale_y_discrete(breaks = c("Other", "Bladder Cancer", "Head and Neck Carcinoma", "Melanoma",
                              "Non-Small Cell Lung Cancer", "Renal Cell Carcinoma", "Sarcoma", "All"), 
                   labels = c("Other", "Bladder Cancer", "Head and Neck\nCarcinoma", "Melanoma",
                              "Non-Small Cell\nLung Cancer", "Renal Cell\nCarcinoma", "Sarcoma", "All")) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        plot.margin = margin(0,0,0,0)) 

```


Combine heatmaps and dendrograms, display and save:

```{r}

test <- grid.arrange(
  grobs = list(dendplot, heatplot),
  widths = c(1.7,2,.06),
  heights = c(1, 4),
  layout_matrix = rbind(c(NA, 1, NA),
                        c(2,2,2))
) 

```


```{r}

test2 <- grid.arrange(
  grobs = list(dbiddendplot, dbidheatplot),
  widths = c(.55,2,.5),
  heights = c(1, 4),
  layout_matrix = rbind(c(NA, 1, NA),
                        c(2,2,2))
) 

```



# Store information on hazard ratios from the antibiotics data

```{r}
hazardrats <- bind_rows(hrbycancbac) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "cancer")

colnames(hazardrats) <- c("cancer", names(hrbycancbac$Other))
colnames(hazardrats)[4] <- 'Linezolid'

hazardrats <- hazardrats %>%
  gather(-cancer, key = "AbxClass", value = "HazardRatio") %>%
  mutate(HazardRatio = ifelse(is.na(HazardRatio), 1, HazardRatio)) %>%
  spread(key = AbxClass, value = HazardRatio) %>%
  column_to_rownames(var = "cancer")
```

```{r}
Median.HR  <- lapply(hazardrats, function(x) median(x, na.rm = TRUE))

Median.HR <- bind_rows(Median.HR)
```

```{r ranks of the classes in cancers}
rankinghrs <- t(hazardrats) %>%
  as.data.frame()

rankinghrs <- lapply(rankinghrs, function(x) as.numeric(as.factor(x)))

rankinghrs <- bind_cols(rankinghrs) %>%
  as.data.frame()

rownames(rankinghrs) <- colnames(hazardrats)
rankinghrs <- t(rankinghrs) %>%
  as.data.frame()

```

```{r}
Times.Worst <- lapply(rankinghrs, function(x) {length(subset(x, x == 1))})

Times.Worst <- bind_rows(Times.Worst)

Times.Worst.3 <- lapply(rankinghrs, function(x) {length(subset(x, x <= 3))})

Times.Worst.3 <- bind_rows(Times.Worst.3)

Average.Rank <- lapply(rankinghrs, function(x) mean(x, na.rm = TRUE))

Average.Rank <- bind_rows(Average.Rank)

sumaryhazardranks <- rbind(Times.Worst = Times.Worst,
                           Times.Worst.3 = Times.Worst.3,
                           Average.Rank = Average.Rank,
                           Median.HR = Median.HR) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column(var = "Antibacterial") %>%
  arrange(Average.Rank)

```

```{r}
saveRDS(sumaryhazardranks, "hazard-table.RDS")
```