---
title: "Dag-causal-model"
author: "Mitchell Muniak"
date: "10/22/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggdag)
library(ggplot2)
library(ggraph)
library(dplyr)

```


# Create dag object for model
+ Each term below will act as a 'node'
+ The dagify function allows for pathways between different nodes
+ For example 'Overall_Survival ~ ICI_Response' means the ICI_Response node leads to the Overall_Survival node
+ Further, 'Prostaglandin ~ Age + ECOG + Staging + BMI + Microbiome' means that Age, ECOG, Staging, BMI, and Microbiome all lead to the Prostaglandin node independently
```{r}

ppi_dag <- dagify(
   Overall_Survival ~ ICI_Response,
   ICI_Response ~ T_cell,
   T_cell ~ Microbiome + Prostaglandin + ECOG + Age,
   Microbiome ~ PPI + Prostaglandin,
   Prostaglandin ~ Age + ECOG + Staging + BMI + Microbiome,
   PPI ~ CCI,
   ECOG ~ CCI + Age,
   BMI ~ ECOG,
   # Indicate the starting node
   exposure = "CCI",
   # Indicate the final node
   outcome = "Overall_Survival",
   # Assign labels to each node for better visualization 
   labels = c("Overall_Survival" = "Overall_Survival",
                   "ICI_Response" = "ICI_Response",
                   "T_cell" = "T_cell",
                   "Microbiome" = "Microbiome",
                   "PPI" = "PPI",
                   "Prostaglandin" = "Prostaglandin",
                   "ECOG" = "ECOG",
                   "BMI" = "BMI",
              "Age" = "Age",
              "CCI" = "CCI",
              "Staging" = "Staging")
)

```

# Plot Dag object in causal model
```{r}
# This will plot the nodes and pathways indicated in the 'ppi_dag' object
# Text over each node is canceled in favor of a label pointed to each node. This allows for a clearer representation.
ggdag(ppi_dag, text = FALSE, use_labels = "label") +
   # The model is saved as a pdf to provide a clearer image
   ggsave("../figures/ppi_dag.pdf", width = 10, height = 10)
```
Pathway visualization
```{r}
# Due to the nature of this model, there  are many different pathways from the exposure to outcome node. The below function color codes each pathway.
ggdag_paths_fan(ppi_dag,
            node_size = 7,
            text_size = 3,
          text = FALSE,
          # Labels used for better visualization
          use_labels = "label",
      spread = 2.5) +
   # The model is saved as a pdf to provide a clearer image
   ggsave("../figures/ppi_dag_col-paths.pdf", width = 10, height = 10)

```
```{r}
# Like the above model, this model shows each pathway present in the 'ppi_dag' object. However, this function plots each pathway as a separate model.
ggdag_paths(ppi_dag,
            node_size = 4,
            text = FALSE,
          # Labels used for better visualization
            use_labels = "label",
            shadow = TRUE,
            text_col = "blue") +
   # The model does not scale well in an RMD or HTML, so I recommend viewing the pdf that I save below
   ggsave("../figures/ppi_dag_full-paths.pdf", width = 20, height = 20)

```