---
title: "abx_class-assign_improve"
author: "Rebecca Hoyd & Mitchell Muniak"
date: "11/23/2020"
output: html_document
---

The purpose of this document is to process the raw data from the warehouse pulls onto the redcap data frame. This document will be split between the actual assignment of medications (first half) and the investigation of different stratification methods etc. (second half). At the time of creation, I have only copied over and edited what code is currently necessary to recreate the processed data used in the further analyses.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("install.load")) {
  install.packages("install.load")
  library(install.load)
}

# CRAN packages
list.of.packages <- c("tidyverse", "glue", "readxl")

install.load::install_load(list.of.packages)
```


```{r load data, include = FALSE}

source("00-paths.R")
# Line where you load the db resulting from format and filter
load(file.path(paths$exoio_raw,"db-meds.RData"))
 # This is from T:/Labs/Spakowicz/data/exoio/00_format-and-filter
db <- read.csv(file.path(paths$exoio_raw, "format_output.csv"))

# This should be in T drive
comorb <-read_excel(file.path(paths$exoio_raw,"CharlsonScore_Presley_08272019 acj.xlsx"), sheet = 2)
# comorb.2 <-read_excel("~/controlled/CharlsonScore_Presley_07152019_Addtl.xlsx")
#steroids <- read_excel("~/controlled/corticosteroid_study678.xlsx", sheet = 2)

# This should be in T drive 
steroids <- readRDS(file.path(paths$exoio_raw,"steroids_validate.RDS"))
sterfun <- read.table(file.path(paths$exoio_raw, "steroid-dbid-key.TSV")) 
bets <- sterfun[1,]
sterfun[1,] <- sterfun[25,]
sterfun[25,] <- bets

stergenkey <- unique(sterfun[c("DBID", "Generic.ID")])

# thromplate <- read_excel("~/controlled/MarAdmin_Thrombin_Inhibitors_study678.xlsx",range = cell_rows(2:30602))

noio <- readRDS("T:/Labs/Spakowicz/projects/ppiio/data/curated/NoImmunotherapy.RDS")

# comorb <- comorb %>%
#   filter(!MRN %in% comorb.2$MRN)
# comorb.2$`TOTAL SCORE` = comorb.2$Total_Charlson
# comorb <- bind_rows(comorb, comorb.2)

ppi_dates <- readRDS(file.path(paths$exoio_raw, "ppi_dates.RDS"))
```


```{r}
comorb <- comorb %>%
  mutate(record_id = as.character(MRN),
         comorb.score = Total) %>%
  select(record_id, comorb.score)


db <- db %>%
  filter(!(record_id %in% noio)) %>%
  left_join(., comorb) %>%
  mutate(othercancer = ifelse(is.na(cancer.name), "Unknown", othercancer),
         cancer.name = ifelse(is.na(cancer.name), "Unknown", cancer.name))

db2 <- db

meds <- meds %>%
  filter(ROUTE != "Apply externally")
```

# Create atb key and assign antibiotic column to meds

Some quick notes on these generic categories:
  Sulfamethoxazole is always prescribed with Trimethoprim
  Piperacillin is always prescribed with Tazobactam
  Imipenem is prescribed with Cilastatin
  
```{r make abx dfs}
blac <- c("penicillin", "Amoxicillin", "Ampicillin", "Piperacillin", "Cefalexin", "Cephalexin", "Ceftriaxone", "Ceftazidime", "Cefepime", "Aztreonam", "Imipenem")
blacname <- rep("BetaLactam", length(blac))
BLactam <- as.data.frame(cbind(blacname, blac))
colnames(BLactam) <- c("Class", "Antibiotic")

FQ <- c("Ciprofloxacin", "levofloxacin", "Moxifloxacin")
FQnames <- rep("Fluoroquinolone", length(FQ))
FQ <- as.data.frame(cbind(FQnames, FQ))
colnames(FQ) <- c("Class", "Antibiotic")

macabx <- c("Azithromycin", "Clarithromycin", "Erythromycin")
macnam <- rep("Macrolide", length(macabx))
macrolides <- as.data.frame(cbind(macnam, macabx))
colnames(macrolides) <- c("Class", "Antibiotic")

tetabx <- c("Doxycycline", "Tetracycline", "Minocycline")
tetnam <- rep("Tetracycline", length(tetabx))
tetracyclines <- as.data.frame(cbind(tetnam, tetabx))
colnames(tetracyclines) <- c("Class", "Antibiotic")

othabx <- c("Linezolid", "Daptomycin")
othnam <- rep("Other.Antibacterial", length(othabx))
others <- as.data.frame(cbind(othnam, othabx))
colnames(others) <- c("Class", "Antibiotic")

antvir <- c("Acyclovir", "Ganciclovir", "Valacyclovir", "Valaciclovir", "Valganciclovir", "Foscarnet", "Oseltamivir")
antvirnam <- rep("Antiviral", length(antvir))
antiviral <- as.data.frame(cbind(antvirnam, antvir))
colnames(antiviral) <- c("Class", "Antibiotic")

antfun <- c("Fluconazole", "Voriconazole", "Posaconazole", "Caspofungin", "Micafungin")
antfunnam <- rep("Antifungal", length(antfun))
antifungals <- as.data.frame(cbind(antfunnam, antfun))
colnames(antifungals) <- c("Class", "Antibiotic")

vanc <- rep("Vancomycin", 2)
trim <- rep("Sulfamethoxazole", 2)
clin <- rep("Clindamycin", 2)
metr <- rep("Metronidazole", 2)
```

```{r combine to key}
abxkey <- as.data.frame(rbind(BLactam, FQ, macrolides, tetracyclines, others, antiviral, antifungals))
abxkey <- bind_cols(lapply(abxkey, function(x) as.character(x)))
abxkey <- as.data.frame(rbind(abxkey, vanc, trim, clin, metr))

abxkey$Class <- as.character(abxkey$Class)
abxkey$Antibiotic <- as.character(abxkey$Antibiotic)
abxkey$organism.type <- ifelse(abxkey$Class == "Antifungal", "Antifungal", 
                       ifelse(abxkey$Class == "Antiviral", "Antiviral", "Antibacterial"))
```

```{r assign classes to meds}
classes <- vector(length = length(meds$GENERIC_NAME))

i <- 5
for(i in 1:length(abxkey$Antibiotic)){
  x <- grep(x = meds$MED_DESCRIPTION, pattern = abxkey$Antibiotic[i], ignore.case = TRUE)
  classes[x] <- abxkey$Antibiotic[i]
  
}
 
check <- as.data.frame(cbind(meds$GENERIC_NAME, classes)) 
y <- grep(check[,2], pattern = "FALSE", ignore.case = TRUE)
check[y,]
classes[y] <- "levofloxacin"
```

```{r}
meds$Antibiotic <- classes
meds <- abxkey %>%
  mutate(abxClass = Class) %>%
  dplyr::select(-Class) %>%
  left_join(meds, .) %>%
  mutate(Antibacterial = ifelse(organism.type == "Antibacterial", 1, 0)) %>%
  mutate(Antiviral = ifelse(organism.type=="Antiviral", 1, 0)) %>%
  mutate(Antifungal = ifelse(organism.type == "Antifungal", 1, 0))

meds <- meds[!duplicated(meds),] 
```

###Break to make list of meds for supplemental table S3 (Is it also a figure now?)
```{r, eval = F}
restrictedmeds <- meds %>%
  filter(days.abx.to.iostart >= -28 & days.abx.to.iostart <= 28)

```

```{r, eval = F}
gen.meds.labs <- c("Acyclovir", "Amoxicillin", "Amoxicillin-Clavulanate", "Ampicillin", "Ampicillin-Sulbactam", "Azithromycin", "Aztreonam", "Caspofungin", "Cefepime", "Ceftriaxone", "Ciprofloxacin", "Clarithromycin", "Clindamycin", "Daptomycin", "Doxycycline", "Erythromycin", "Fluconazole", "Levofloxacin", "Linezolid", "Metronidazole", "Minocycline", "Oseltamivir", "Penicillin-G", "Penicillin-V", "Piperacillin-Tazobactam", "Sulfamethoxazole-Trimethoprim", "Valacyclovir", "Vancomycin", "Voriconazole", "Cephalexin", "Moxifloxacin")

gen.meds.pat <- c("Acyclovir", "Amoxicillin", "Clavulanate", "Ampicillin", "Sulbactam", "Azithromycin", "Aztreonam", "Caspofungin", "Cefepime", "Ceftriaxone", "Ciprofloxacin", "Clarithromycin", "Clindamycin", "Daptomycin", "Doxycycline", "Erythromycin", "Fluconazole", "Levofloxacin", "Linezolid", "Metronidazole", "Minocycline", "Oseltamivir", "Penicillin G", "Penicillin V", "Piperacillin", "Sulfamethoxazole", "Valacyclovir", "Vancomycin", "Voriconazole", "Cephalexin", "Moxifloxacin")


classes <- vector(length = length(restrictedmeds$GENERIC_NAME))

i <- 5
for(i in 1:length(gen.meds.pat)){
  x <- grep(x = restrictedmeds$GENERIC_NAME, pattern = gen.meds.pat[i], ignore.case = TRUE)
  classes[x] <- gen.meds.labs[i]
  
}
 
check <- as.data.frame(cbind(restrictedmeds$GENERIC_NAME, classes)) 
y <- grep(check[,2], pattern = "FALSE", ignore.case = TRUE)
check[y,]
```


### save info on other antibiotics


```{r for github other characterize}
oth.med.break <- meds %>%
  filter(days.abx.to.iostart >= -28 & days.abx.to.iostart <= 28) %>%
  filter(abxClass == "Other.Antibacterial") %>%
  group_by(MRN) %>%
  summarise(oth.presc.28 = glue_collapse(unique(Antibiotic), sep = ","))

db.oth.med.break <- db %>%
  filter(record_id %in% oth.med.break$MRN) %>%
  mutate(MRN = as.numeric(record_id)) %>%
  left_join(oth.med.break) %>%
  select(record_id, oth.presc.28)
```


```{r create some functions for column assignment}
binarize <- function(vect){
  ifelse(sum(vect, na.rm = TRUE) >= 1, 1, 0)
}

narem <- function(vect){
  ifelse(is.na(vect), 0, vect)
}
```

## Adding timing information
```{r}
abxdaysstring <- meds %>%
  mutate(record_id = MRN) %>%
  filter(days.abx.to.iostart >= -180 & days.abx.to.iostart <= 180) %>%
  select(-MRN) %>%
  group_by(record_id) %>%
  summarise(abxdays.to.iostart = glue_collapse(unique(days.abx.to.iostart), sep = ","))

abxdaysstring$record_id <- as.character(abxdaysstring$record_id)

db <- db %>%
  left_join(abxdaysstring)

db2 <- db2 %>%
  left_join(abxdaysstring)
```


### Add timing info per class to alternate db
```{r}
abx.all.days <- meds %>%
  mutate(record_id = MRN,
         record_id = as.character(record_id)) %>%
  filter(days.abx.to.iostart >= -180 & days.abx.to.iostart <= 180) %>%
  mutate(organism.type = paste(organism.type, ".to.start", sep = ""),
         abxClass = paste(abxClass, ".to.start", sep = ""))

abx.org.days <- abx.all.days %>%
  group_by(record_id, organism.type) %>%
  summarise(days.to.start = glue_collapse(unique(days.abx.to.iostart), sep = ",")) %>%
  spread(key = "organism.type", value = "days.to.start")

abx.class.days <- abx.all.days %>%
  group_by(record_id, abxClass) %>%
  summarise(days.to.start = glue_collapse(unique(days.abx.to.iostart), sep = ",")) %>%
  spread(key = "abxClass", value = "days.to.start") %>%
  select(-Antifungal.to.start, -Antiviral.to.start)


db2 <- db2 %>%
  left_join(abx.org.days) %>%
  left_join(abx.class.days)
```

## Adding abx classes, with window considerations
```{r anti classes window 60 before and after,eval=FALSE}

abx.60.60meds <- meds %>%
  filter(abx.60.60 == 1)

abx.60.60meds <- abx.60.60meds %>%
  mutate(BetaLactam = ifelse(abxClass == "BetaLactam", 1, 0))%>%
  mutate(Macrolide = ifelse(abxClass == "Macrolide", 1, 0))%>%
  mutate(Other.Antibacterial = ifelse(Antibiotic =="Sulfamethoxazole"| Antibiotic == "Linezolid"| Antibiotic == "Daptomycin", 1, 0))%>%
  mutate(Fluoroquinolone = ifelse(abxClass == "FQ", 1, 0))%>%
  mutate(Tetracycline = ifelse(abxClass == "Tetracycline", 1, 0))%>%
  mutate(Acyclovir = ifelse(Antibiotic == "Acyclovir", 1, 0)) %>%
  mutate(Valacyclovir = ifelse(Antibiotic == "Valacyclovir", 1, 0)) %>%
  mutate(Metronidazole = ifelse(Antibiotic == "Metronidazole",1,0)) %>%
  mutate(Vancomycin = ifelse(Antibiotic == "Vancomycin", 1, 0)) %>%
  mutate(Sulfamethoxazole = ifelse(Antibiotic == "Sulfamethoxazole", 1, 0)) %>%
  mutate(Clindamycin = ifelse(Antibiotic == "Clindamycin", 1, 0)) %>%
  mutate(Antibacterial = ifelse(organism.type == "Antibacterial", 1, 0)) %>%
  mutate(Antifungal = ifelse(organism.type == "Antifungal", 1, 0)) %>%
  mutate(Antiviral = ifelse(organism.type == "Antiviral", 1, 0)) %>%
  dplyr::select(record_id, BetaLactam, Macrolide, Other.Antibacterial, Fluoroquinolone, Vancomycin, Sulfamethoxazole, Clindamycin, Metronidazole, Tetracycline, Antibiotic, Antibacterial, Antiviral, Antifungal, Acyclovir, Valacyclovir, abxClass)

abxforjoin <- abx.60.60meds %>%
  group_by(record_id) %>%
  summarise(BetaLactam = binarize(BetaLactam),
            Macrolide = binarize(Macrolide),
            Fluoroquinolone = binarize(Fluoroquinolone),
            Tetracycline = binarize(Tetracycline), 
            Other.Antibacterial = binarize(Other.Antibacterial),
            Vancomycin = binarize(Vancomycin),
            Sulfamethoxazole = binarize(Sulfamethoxazole),
            Clindamycin = binarize(Clindamycin),
            Metronidazole = binarize(Metronidazole),
            Antibacterial = binarize(Antibacterial),
            Antiviral = binarize(Antiviral),
            Antifungal = binarize(Antifungal),
            Acyclovir = binarize(Acyclovir),
            Valacyclovir = binarize(Valacyclovir)
            )

db <- db %>%
  left_join(abxforjoin) %>%
  ungroup() %>%
  mutate(
    BetaLactam = narem(BetaLactam),
    Macrolide = narem(Macrolide),
    Fluoroquinolone = narem(Fluoroquinolone),
    Tetracycline = narem(Tetracycline), 
    Other.Antibacterial = narem(Other.Antibacterial),
    Vancomycin = narem(Vancomycin),
    Sulfamethoxazole = narem(Sulfamethoxazole),
    Clindamycin = narem(Clindamycin),
    Metronidazole = narem(Metronidazole),
    Antibacterial = narem(Antibacterial),
    Antiviral = narem(Antiviral),
    Antifungal = narem(Antifungal),
    Acyclovir = narem(Acyclovir),
    Valacyclovir = narem(Valacyclovir)
  )
```

```{r atb window 60 days post, eval = FALSE}
meds$abx.plus.60 <- ifelse(meds$days.abx.to.iostart >= -0 & meds$days.abx.to.iostart <= 60, 1, 0)

abxplus.60meds <- meds %>%
  filter(abx.plus.60 == 1) 

abxplus.60Class <- unique(abxplus.60meds$abxClass)
abxplus.60meds <- abxplus.60meds %>%
  mutate(BetaLactamplus = ifelse(abxClass == "BetaLactam", 1, 0))%>%
  mutate(Macrolideplus = ifelse(abxClass == "Macrolide", 1, 0))%>%
  mutate(Other.Antibacterialplus = ifelse(Antibiotic =="Sulfamethoxazole"| Antibiotic == "Linezolid"| Antibiotic == "Daptomycin", 1, 0))%>%
  mutate(Fluoroquinoloneplus = ifelse(abxClass == "FQ", 1, 0))%>%
  mutate(Tetracyclineplus = ifelse(abxClass == "Tetracycline", 1, 0))%>%
  mutate(Acyclovirplus = ifelse(Antibiotic == "Acyclovir", 1, 0)) %>%
  mutate(Valacyclovirplus = ifelse(Antibiotic == "Valacyclovir", 1, 0)) %>%
  mutate(Metronidazoleplus = ifelse(Antibiotic == "Metronidazole",1,0)) %>%
  mutate(Vancomycinplus = ifelse(Antibiotic == "Vancomycin", 1, 0)) %>%
  mutate(Sulfamethoxazoleplus = ifelse(Antibiotic == "Sulfamethoxazole", 1, 0)) %>%
  mutate(Clindamycinplus = ifelse(Antibiotic == "Clindamycin", 1, 0)) %>%
  mutate(Antibacterialplus = ifelse(organism.type == "Antibacterial", 1, 0)) %>%
  mutate(Antifungalplus = ifelse(organism.type == "Antifungal", 1, 0)) %>%
  mutate(Antiviralplus = ifelse(organism.type == "Antiviral", 1, 0)) %>%
  dplyr::select(record_id, BetaLactamplus, Macrolideplus, Other.Antibacterialplus, Fluoroquinoloneplus, Vancomycinplus, Sulfamethoxazoleplus, Clindamycinplus, Metronidazoleplus, Tetracyclineplus, Antibacterialplus, Antiviralplus, Antifungalplus, Acyclovirplus, Valacyclovirplus, abxClass)

abxplusforjoin <- abxplus.60meds %>%
  group_by(record_id) %>%
  summarise(BetaLactamplus = binarize(BetaLactamplus),
            Macrolideplus = binarize(Macrolideplus),
            Fluoroquinoloneplus = binarize(Fluoroquinoloneplus),
            Tetracyclineplus = binarize(Tetracyclineplus), 
            Other.Antibacterialplus = binarize(Other.Antibacterialplus),
            Vancomycinplus = binarize(Vancomycinplus),
            Sulfamethoxazoleplus = binarize(Sulfamethoxazoleplus),
            Clindamycinplus = binarize(Clindamycinplus),
            Metronidazoleplus = binarize(Metronidazoleplus),
            Antibacterialplus = binarize(Antibacterialplus),
            Antiviralplus = binarize(Antiviralplus),
            Antifungalplus = binarize(Antifungalplus),
            Acyclovirplus = binarize(Acyclovirplus),
            Valacyclovirplus = binarize(Valacyclovirplus)
            )

db <- db %>%
  left_join(abxplusforjoin) %>%
  ungroup() %>%
  mutate(BetaLactamplus = narem(BetaLactamplus),
         Macrolideplus = narem(Macrolideplus),
         Fluoroquinoloneplus = narem(Fluoroquinoloneplus),
         Tetracyclineplus = narem(Tetracyclineplus), 
         Other.Antibacterialplus = narem(Other.Antibacterialplus),
         Vancomycinplus = narem(Vancomycinplus),
         Sulfamethoxazoleplus = narem(Sulfamethoxazoleplus),
         Clindamycinplus = narem(Clindamycinplus),
         Metronidazoleplus = narem(Metronidazoleplus),
         Antibacterialplus = narem(Antibacterialplus),
         Antiviralplus = narem(Antiviralplus),
         Antifungalplus = narem(Antifungalplus),
         Acyclovirplus = narem(Acyclovirplus),
         Valacyclovirplus = narem(Valacyclovirplus))

```


```{r abx classes 60 days before, eval = FALSE}
abxless.60meds <- meds %>%
  filter(abx.60 == 1) 

abxless.60meds <- abxless.60meds %>%
  mutate(BetaLactamless = ifelse(abxClass == "BetaLactam", 1, 0))%>%
  mutate(Macrolideless = ifelse(abxClass == "Macrolide", 1, 0))%>%
  mutate(Other.Antibacterialless = ifelse(Antibiotic =="Sulfamethoxazole"| Antibiotic == "Linezolid"| Antibiotic == "Daptomycin", 1, 0))%>%
  mutate(Fluoroquinoloneless = ifelse(abxClass == "FQ", 1, 0))%>%
  mutate(Tetracyclineless = ifelse(abxClass == "Tetracycline", 1, 0))%>%
  mutate(Acyclovirless = ifelse(Antibiotic == "Acyclovir", 1, 0)) %>%
  mutate(Valacyclovirless = ifelse(Antibiotic == "Valacyclovir", 1, 0)) %>%
  mutate(Metronidazoleless = ifelse(Antibiotic == "Metronidazole",1,0)) %>%
  mutate(Vancomycinless = ifelse(Antibiotic == "Vancomycin", 1, 0)) %>%
  mutate(Sulfamethoxazoleless = ifelse(Antibiotic == "Sulfamethoxazole", 1, 0)) %>%
  mutate(Clindamycinless = ifelse(Antibiotic == "Clindamycin", 1, 0)) %>%
  mutate(Antibacterialless = ifelse(organism.type == "Antibacterial", 1, 0)) %>%
  mutate(Antifungalless = ifelse(organism.type == "Antifungal", 1, 0)) %>%
  mutate(Antiviralless = ifelse(organism.type == "Antiviral", 1, 0)) %>%
  dplyr::select(record_id, BetaLactamless, Macrolideless, Other.Antibacterialless, Fluoroquinoloneless, Vancomycinless, Sulfamethoxazoleless, Clindamycinless, Metronidazoleless, Tetracyclineless, Antibiotic, Antibacterialless, Antiviralless, Antifungalless, Acyclovirless, Valacyclovirless, abxClass)

abxlessforjoin <- abxless.60meds %>%
  group_by(record_id) %>%
  summarise(BetaLactamless = binarize(BetaLactamless),
            Macrolideless = binarize(Macrolideless),
            Fluoroquinoloneless = binarize(Fluoroquinoloneless),
            Tetracyclineless = binarize(Tetracyclineless), 
            Other.Antibacterialless = binarize(Other.Antibacterialless),
            Vancomycinless = binarize(Vancomycinless),
            Sulfamethoxazoleless = binarize(Sulfamethoxazoleless),
            Clindamycinless = binarize(Clindamycinless),
            Metronidazoleless = binarize(Metronidazoleless),
            Antibacterialless = binarize(Antibacterialless),
            Antiviralless = binarize(Antiviralless),
            Antifungalless = binarize(Antifungalless),
            Acyclovirless = binarize(Acyclovirless),
            Valacyclovirless = binarize(Valacyclovirless)
            )

db <- db %>%
  left_join(abxlessforjoin) %>%
  ungroup() %>%
  mutate(BetaLactamless = narem(BetaLactamless),
            Macrolideless = narem(Macrolideless),
            Fluoroquinoloneless = narem(Fluoroquinoloneless),
            Tetracyclineless = narem(Tetracyclineless), 
            Other.Antibacterialless = narem(Other.Antibacterialless),
            Vancomycinless = narem(Vancomycinless),
            Sulfamethoxazoleless = narem(Sulfamethoxazoleless),
            Clindamycinless = narem(Clindamycinless),
            Metronidazoleless = narem(Metronidazoleless),
            Antibacterialless = narem(Antibacterialless),
            Antiviralless = narem(Antiviralless),
            Antifungalless = narem(Antifungalless),
            Acyclovirless = narem(Acyclovirless),
            Valacyclovirless = narem(Valacyclovirless))
```


```{r abx within 60 days administration route, eval = FALSE}
admins <- unique(meds$ROUTE)[-1]
admins

meds.admins <- meds %>%
  filter(abx.60.60 == 1) %>%
  mutate(Oral = ifelse(ROUTE == "Oral", 1, 0),
         Intravenous = ifelse(ROUTE == "Intravenous", 1, 0),
         Otic = ifelse(ROUTE == "Otic", 1, 0),
         Injection = ifelse(ROUTE == "Injection", 1, 0),
         Ophthalmic = ifelse(ROUTE == "Ophthalmic", 1, 0),
         External = ifelse(ROUTE == "Apply externally", 1, 0),
         Unknown = ifelse(ROUTE == "Unknown", 1, 0),
         Vaginal = ifelse(ROUTE == "Vaginal", 1, 0),
         Intramuscular = ifelse(ROUTE == "Intramuscular", 1, 0))

meds.admins <- meds.admins %>%
  group_by(MRN) %>%
  summarise(Oral = binarize(Oral),
         Intravenous = binarize(Intravenous),
         Otic = binarize(Otic),
         Injection = binarize(Injection),
         Ophthalmic = binarize(Ophthalmic),
         External = binarize(External),
         Unknown = binarize(Unknown),
         Vaginal = binarize(Vaginal),
         Intramuscular = binarize(Intramuscular)) %>%
  mutate(record_id = as.character(MRN)) %>%
  select(-MRN)

db <- db %>%
  left_join(meds.admins) %>%
  mutate(Oral = narem(Oral),
         Intravenous = narem(Intravenous),
         Otic = narem(Otic),
         Injection = narem(Injection),
         Ophthalmic = narem(Ophthalmic),
         External = narem(External),
         Unknown.ATB.Route = narem(Unknown),
         Vaginal = narem(Vaginal),
         Intramuscular = narem(Intramuscular)) %>%
  dplyr::select(-Unknown)
          
```

# Steroid assignments

## Calculate days until iostart and add column of all days steroids were prescribed

```{r}
iostart <- db %>%
  select(record_id, iostart) %>%
  mutate(iostart = as.Date(iostart, format = "%F")) %>%
  mutate(MRN = record_id) %>%
  select(-record_id)

steroids <- steroids %>%
  mutate(MRN = as.character(MRN)) %>%
  left_join(iostart)

steroids$iostart <- as.Date(steroids$iostart, format = "%F")
steroids$START_DATE <- as.Date(steroids$START_DATE, formate = "%F")
steroids$days.to.iostart <- steroids$START_DATE - steroids$iostart

steroids <- steroids %>% 
  filter(!is.na(days.to.iostart)) %>%
  mutate(ster.30 = ifelse(days.to.iostart > -30 & days.to.iostart <= 0, 1, 0)) %>%
  mutate(ster.30.30 = ifelse(days.to.iostart > -30 & days.to.iostart <= 30, 1, 0)) %>%
  mutate(ster.60 = ifelse(days.to.iostart > -60 & days.to.iostart <= 0, 1, 0)) %>%
  mutate(ster.60.60 = ifelse(days.to.iostart > -60 & days.to.iostart <= 60, 1, 0)) %>%
  mutate(ster.180.180 = ifelse(days.to.iostart > -180 & days.to.iostart <= 180, 1, 0))

daysstring <- steroids %>%
  mutate(record_id = MRN) %>%
  filter(days.to.iostart >= -180 & days.to.iostart <= 180) %>%
  select(-MRN) %>%
  group_by(record_id) %>%
  summarise(sterdays.to.start = glue_collapse(unique(days.to.iostart), sep = ","))


db2 <- db2 %>%
  left_join(daysstring)
```

## Assing route by which the steroids were administered

```{r create steroid application key}
sterroute <- c("Injection", "Oral", "Apply externally", "Rectal", "Ophthalmic", "Unknown", "Otic" ) 

sterclass <- c("IV", "Oral", rep("Other.route", 5))

sterassign <- as.data.frame(cbind(sterroute, sterclass)) %>%
  mutate(sterroute = as.character(sterroute),
         sterclass = as.character(sterclass))
```

```{r}
classes <- vector(length = length(steroids$GENERIC_NAME))

for(i in 1:length(sterroute)){
  x <- grep(x = steroids$ROUTE, pattern = sterroute[i], ignore.case = TRUE)
  classes[x] <- sterassign[i, 2]
  
}
 
check <- as.data.frame(cbind(steroids$ROUTE, classes)) 
y <- grep(check[,2], pattern = "FALSE", ignore.case = TRUE)
check[y, "classes"] <- NA
```

## Assign steroid class generic names
```{r}
sterforfun <- vector(length = length(steroids$GENERIC_NAME))
sterdrugname <- sterfun$Drug.name

i <- 5
for(i in 1:length(sterdrugname)){
  x <- grep(x = steroids$GENERIC_NAME, pattern = sterdrugname[i], ignore.case = TRUE)
  sterforfun[x] <- as.character(sterfun[i,1])
  
}
 
check <- as.data.frame(cbind(steroids$GENERIC_NAME, sterforfun)) 
y <- grep(check[,2], pattern = "FALSE", ignore.case = TRUE)
check[y,]
```

```{r, eval = FALSE}
steroids$sterclass <- classes
steroids$sterforfun <- sterforfun
```

### Add timing info for classes
```{r}
ster.class.time <- steroids %>%
  mutate(record_id = MRN,
         DBID = sterforfun) %>%
  select(-MRN) %>%
  filter(ster.180.180 == 1) %>%
  left_join(stergenkey) %>%
  mutate(Generic.ID = paste(Generic.ID, ".to.start", sep = "")) %>%
  group_by(record_id, Generic.ID) %>%
  summarise(days.to.start = glue_collapse(unique(days.to.iostart), sep = ",")) %>%
  spread(key = "Generic.ID", value = "days.to.start")

db2 <- db2 %>%
  left_join(ster.class.time)
```


```{r}
# Create column for time of irae relevant to iostart
# This was created as its own data frame because not every subject has iostart and date_of_irae data. Filtering to subjects with that data was required to calculate irae_time_to_start below. I found it easier to do this as its own data frame, then join it back to db2 via record_id
irae_time <- db2 %>%
  # Clear empty rows
  filter(iostart != "is.na", date_of_irae != "") %>%
  # Represent as days after iostart
  mutate(irae_time_to_start = as.Date(date_of_irae) - as.Date(iostart)) %>%
  # Keep calculated column and record_id so it can be joined to db2
  select(record_id, irae_time_to_start)
```


```{r}
# ppi_iostart <- db2 %>%
#   filter(!is.na(ppi_start) & ppi_start != "") %>%
#   mutate(ppidays.to.iostart = as.Date(ppi_start) - as.Date(iostart)) %>%
#   mutate(ppidays.to.iostart = gsub(" days", "", ppidays.to.iostart)) %>%
#   select(ppidays.to.iostart, record_id)
```


```{r}
deidentdb2 <- db2 %>%
  left_join(db.oth.med.break) %>%
  left_join(irae_time) %>%
  left_join(ppi_dates) %>%
  # mutate(race = as.character(race),
  #        vitalstatus = as.character(vitalstatus),
  #        ecog = as.character(ecog),
  #        comorb.score = as.character(comorb.score),
  #        cancer = as.character(cancer),
  #        staging = as.character(staging)) %>%
  dplyr::select(X, race, sex, age, bmi.calc, vitalstatus, ecog, comorb.score, cancer, cancer.name, staging, abxdays.to.iostart, sterdays.to.start, ppidays.to.iostart.1, ppidays.to.iostart.2, ppidays.to.iostart.3, days)
```

#Saving data
```{r}
# I can save this to the t drive
saveRDS(deidentdb2, file.path(paths$ppiio_derived, "deidentdb_timing.RDS"))

```
