---
title: 'Figure 3: combined model'
author: "Rebecca Hoyd"
date: "May 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(survminer)
library(survival)
library(survMisc)
library(ggplot2)
library(dplyr)
library(glue)
library(glmnet)
```


```{r}

db <- readRDS("../data/deidentdb_timing.RDS") 

# purge github cache
# Remove dates 
# List out ppi columns in the above data frame
# Move processing to abx_class_improve
# Find pdf that has column des

```

```{r}
check.window <- function(vect, per){
  wind <- per[1]:per[2]
  wind <- glue_collapse(as.character(wind), sep = ",|,")
  wind <- paste(",", wind, ",", sep = "")
  ifelse(grepl(pattern = wind, x = vect), 1, 0)
}

db <- db %>%
  mutate(ppidays.to.iostart = paste(",", ppidays.to.iostart, ",", sep = ""),
         abxdays.to.iostart = paste(",", abxdays.to.iostart, ",", sep = ""),
         sterdays.to.start = paste(",", sterdays.to.start, ",", sep = ""),
         ppi.90.90 = check.window(ppidays.to.iostart, c(-90,90)),
         abx.28.28 = check.window(abxdays.to.iostart, c(-28,28)),
         steroid.28.28 = check.window(sterdays.to.start, c(-28,28)))
```

# A) Survival curves
All meds
+Write formula
```{r}
sfit_all <- survfit(Surv(days, vitalstatus) ~ abx.28.28 + steroid.28.28 + ppi.90.90, data = db)
```

Create survival curve 
```{r}

comb_curv_all <- ggsurvplot(fit = sfit_all, conf.int = FALSE, risk.table = TRUE, tables.height = 0.45, pval = TRUE, ylab = "", 
                        legend = "none", legend.title = "", 
                        legend.labs = c("ABx, CS, PPI=0", "ABx, CS=0; PPI=1",
                                        "ABx, PPI=0;CS=1", "ABx=0; CS, PPI = 1",
                                        "ABx=1; CS, PPI=0", "ABx, PPI=1; CS=0",
                                        "ABx, CS=1; PPI=0", "ABx, CS, PPI=1"),
                        font.legend = c(6), fontsize = 4, xlab = "Days", pval.coord =
                          c(0,0.08), 
                        palette = "Dark2")

comb_curv_all
```

```{r}
comb_curv_all.leg <- ggsurvplot(fit = sfit_all, conf.int = FALSE, risk.table = TRUE, tables.height = 0.45, pval = TRUE,
                            legend = "right", legend.title = "", 
                            legend.labs = c("ABx, CS, PPI=0", "ABx, CS=0; PPI=1", 
                                        "ABx, PPI=0;CS=1", "ABx=0; CS, PPI = 1",
                                        "ABx=1; CS, PPI=0", "ABx, PPI=1; CS=0",
                                        "ABx, CS=1; PPI=0", "ABx, CS, PPI=1"), 
                            font.legend = c(8), fontsize = 4, xlab = "Days",
                            pval.coord = c(0,0.08), palette = "Dark2")

comb_curv_all.leg
```
PPIs only
```{r}
sfit_ppi <- survfit(Surv(days, vitalstatus) ~ ppi.90.90, data = db)

comb_curv_ppi <- ggsurvplot(fit = sfit_ppi, conf.int = FALSE, risk.table = TRUE, tables.height = 0.28, pval = TRUE, ylab = "", 
                            legend = "none", legend.title = "", 
                            legend.labs = c("No PPI -90 Days", "Yes PPI -90 Days"),
                            font.legend = c(6), fontsize = 4, xlab = "Days", pval.coord = c(0,0.08), 
                            palette = "Dark2")

comb_curv_ppi
```

```{r}
comb_curv_ppi.leg <- ggsurvplot(fit = sfit_ppi, conf.int = FALSE, risk.table = TRUE,  tables.height = 0.30, pval = TRUE,
                            legend = "right", legend.title = "", 
                            legend.labs = c("No PPI -90 Days", "Yes PPI -90 Days"), 
                            font.legend = c(8), fontsize = 4, xlab = "Days",
                            pval.coord = c(0,0.08), palette = "Dark2")

comb_curv_ppi.leg
```

# B) Combined model for all cancers

```{r}

# Add as.factor
db <- db %>%
  mutate(ABx = abx.28.28,
         CS = steroid.28.28,
         PPI = ppi.90.90,
         ECOG = ecog,
         CCI = ifelse(comorb.score <= 1, 0, 1),
         BMI = bmi.calc,
         Sex = sex,
         Age = age,
         `Stage IV` = ifelse(staging == 4, 1, 0))

cph <- coxph(Surv(days, vitalstatus) ~ ABx + CS + PPI + ECOG + CCI + `Stage IV` + BMI + Sex + Age, data = db)

ggforest(cph) 

```


# Loop performing percentile lasso to generate data for remaining panels

Please try to avoid unnecessarily rerunning the followingcode, it take ~20 minutes to complete this loop.
```{r perform percentile lasso takes a long time, eval = F}
# allcancers <- c('Other', 'Melanoma', 'Non-Small Cell Lung Cancer', 'Head and Neck Carcinoma', 'Renal Cell Carcinoma', 'Sarcoma', 'Bladder Cancer')



allcancerspre <- db %>%
  filter(!is.na(ECOG) & !is.na(BMI) & !is.na(`Stage IV`) & !is.na(Age) & 
             !is.na(CCI))

allcancers <- unique(allcancerspre$cancer.name)
allcancers <- allcancers[-11]


all.percentile.res <- list()
canc.percentile.res <- list()

box.res <- list()
canc.box.res <- list()



# c <- "melanoma"

for(c in allcancers){
  print(c)
  tmp <- db %>%
    filter(!is.na(ECOG) & !is.na(BMI) & !is.na(`Stage IV`) & !is.na(Age) & 
             !is.na(CCI)) %>%
    filter(cancer.name == c)
  
  x <- tmp %>%
    mutate(Male = ifelse(Sex == "Male", 1, 0)) %>%
    select(PPI, ABx, CS, ECOG, CCI, BMI, Age, `Stage IV`, Male) %>%
    as.matrix()
  
  set.seed(112358)
  fit <- glmnet(x, Surv(tmp$days, tmp$vitalstatus), family = "cox", maxit = 1000)
  for(i in 1:1000){
    try({
    cv.fit <- cv.glmnet(x, y = Surv(tmp$days, tmp$vitalstatus), family="cox", maxit =
                          1000)
    
    Coefficients <-  coef(fit, s = cv.fit$lambda.min)
    Active.Index <- which(Coefficients != 0)
    
    canc.percentile.res[[i]] <- as.data.frame(cbind(cancer = c,
                                                    lambdahat = cv.fit$lambda.min,
                                                    covcount = length(Active.Index),
                                                    whichco=glue_collapse(Coefficients@Dimnames[[1]][Active.Index],
                                                                            sep = "+")))
    
    canc.box.res[[i]] <- as.data.frame(cbind(term = Coefficients@Dimnames[[1]][Active.Index], 
                                        HR = exp(Coefficients[Active.Index]),
                                        cancer = c))
    })
  }
  
  all.percentile.res[[c]] <- bind_rows(canc.percentile.res)
  canc.percentile.res <- list()
  
  box.res[[c]] <- bind_rows(canc.box.res)
  canc.box.res <- list()
}

lassores <- bind_rows(all.percentile.res)
lassores$lambdahat <- as.numeric(lassores$lambdahat)
lassores$covcount <- as.factor(lassores$covcount) 

# saveRDS(lassores, "fig4_percentile-lasso.RDS")
# saveRDS(box.res, "fig4-sup_percent-appearance-and-hr_boxplots")

```

```{r}
# lassores <- readRDS("fig4_percentile-lasso.RDS")
# box.res <- readRDS("fig4-sup_percent-appearance-and-hr_boxplots.RDS")
```

```{r}
box.res <- bind_rows(box.res)
```

# C) Boxplots of HR results from lasso with barplots of how frequently models selecting given covariates were chosen

```{r}
box.res$HR <- as.numeric(box.res$HR)

box.res  %>%
  filter(!is.na(term)) %>%
  mutate(cancer = ifelse(cancer == "Head and Neck Carcinoma", "Head and Neck \nCarcinoma", 
                         ifelse(cancer == "Non-Small Cell Lung Cancer", "Non-Small Cell \nLung Cancer", 
                                ifelse(cancer == "Renal Cell Carcinoma", "Renal Cell \nCarcinoma", cancer)))) %>%
  ggplot(aes(x = term, y = HR, fill = term)) +
  facet_wrap(vars(cancer), strip.position = "left", scales = "free", nrow = 9) + 
  geom_boxplot(alpha = .4) +
  geom_point(aes(color = term)) +
  geom_hline(aes(fill = term), yintercept = 1, linetype =2)+
  scale_color_brewer(aesthetics = c("color", "fill"), palette = "Dark2") +
  theme_bw() +
  theme(plot.title=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_blank(),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold", size = 8)) +
  coord_flip() 
```


```{r}
box.res %>%
  filter(!is.na(term)) %>%
  ggplot(aes(x = term, fill = term)) + 
  geom_bar() +
  facet_wrap(vars(cancer), scales = "free_y", strip.position = "left",nrow = 9) +
  scale_color_brewer(aesthetics = c("color", "fill"), palette = "Dark2", name = "Covariate") +
  labs(x = "", y = "") +
  theme_bw() +
  theme(plot.title=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x=element_text(face="bold"),
        axis.title=element_blank(),
        strip.text.y = element_text(hjust=0,vjust = 1,angle=180,face="bold"),
        panel.spacing = unit(1.2, "lines")) +
  coord_flip()+
  guides(fill = guide_legend(reverse = TRUE)) 
```