---
title: "medication-summary_orien"
author: "Dan Spakowicz"
date: "2/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

this_dir <- if (!is.null(knitr::current_input())) {
  dirname(normalizePath(knitr::current_input()))
} else {
  getwd()
}
# Older notebooks used a local 00-paths.R with a `paths` list (e.g. paths$phi).
local_paths <- file.path(this_dir, "00-paths.R")
if (file.exists(local_paths)) source(local_paths)

# Tdrive is the ppiio folder on the lab share (gitignored 00-paths.R).
paths_file <- file.path(this_dir, "..", "..", "manuscript", "scripts", "00-paths.R")
if (!file.exists(paths_file)) {
  stop(
    "Missing 00-paths.R. Copy manuscript/scripts/00-paths.R.example to ",
    "00-paths.R and set Tdrive to the ppiio folder on the lab share."
  )
}
source(paths_file)
if (!exists("Tdrive") || !nzchar(Tdrive)) {
  stop("00-paths.R must define Tdrive as the ppiio folder on the lab share.")
}
```

This is a summary of medication data shared with Dan on Feb 22, 2021 by Rachel 

```{r}
files <- list.files(file.path(paths$phi, "TCC000139"),
                    full.names = TRUE)
files
```

```{r}
x <- readxl::read_excel(files[grep("ClinicalDataPoints_withSLIDS", 
                                   files
                                   )
                              ]
                        )

dim(x)
names(x)
table(x$Tissue_Type_RNA_Seq)
```

Why 686? Are these the files that are also in Dwight's database? Not sure why the tissue type was included.

```{r}
y <- readxl::read_excel(files[grep("MedicationLogFile_withSLIDS", files)],
                        na = "NULL") %>%
  mutate(Simple_Generic_Name = if_else(is.na(Simple_Generic_Name),
                                       true = Clarity_Med_Name,
                                       false = Simple_Generic_Name))

dim(y)
names(y)
```

```{r}
y %>%
  select(`RNA-SLID`) %>%
  unique() %>%
  tally()
```

```{r}
  y %>%
  group_by(`RNA-SLID`) %>%
  summarize(meds.per.pt = n()) %>%
  ggplot(aes(x = meds.per.pt)) +
  geom_histogram() +
  theme_bw()

```

```{r}
y %>%
  ggplot(aes(x = Days_Btwn_Collect_and_Med_Start)) +
  geom_histogram() +
  theme_bw()
```

```{r}
y %>%
  group_by(Simple_Generic_Name) %>%
  summarize(n = n()) %>%
  arrange(-n) %>%
  top_n(20) %>%
  ggplot(aes(x = reorder(Simple_Generic_Name, -n), y = n)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("")
```

How many SLIDs are in CoRR4684?

```{r}
z <- readxl::read_excel(files[grep("patientsRNAseqAvail", files)])

dim(z)
names(z)
```

```{r}
sum(!is.na(z$`TCC SL ID`))

sum(z$`TCC SL ID` %in% x$`RNA-SLID`)
```

Well that's not very many.

Check if we have RNAseq data for these in PAS1695. To facilitate this, output a list of the SLIDs for which we have med information to compare to other lists.

```{r}
slids <- 
  y %>%
  select(`RNA-SLID`) %>%
  unique

write.csv(slids,
          file = file.path(Tdrive, "data/slids.csv"),
          row.names = FALSE)
```

