---
title: "table_one-ppi"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(tidyverse)
library(tableone)
library(forcats)
library(glue)
library(eeptools)
library(readxl)
```


```{r}
cancer.copy <- c("1 non-small cell lung cancer
2 small cell lung cancer
3 melanoma
4 renal cell carcinoma
5 head and neck carcinoma
6 merkel cell carcinoma
7 Hodgkin lymphoma
8 breast cancer
9 colon cancer
10 pancreatic cancer
11 Sarcoma
12 Prostate
13 bladder cancer
14 AML
15 ALL
17 Other") %>%
  strsplit(., split = "\n") %>%
  unlist

cancer.key <- 
  data.frame(
    cancer =  cancer.copy %>%
      gsub("(\\d+).*", "\\1", .) %>%
      as.integer(),
    cancer.name = cancer.copy %>%
      gsub("\\d+ (.*)", "\\1", .),
    stringsAsFactors = FALSE
  )
```

# load data
```{r}
db <- readRDS("T:/Labs/Spakowicz/ppiio/data/raw/table_one_db.RDS") # want this to be in ppiio
CoRR <- read.csv("T:/Labs/Spakowicz/ppiio/data/raw/CoRR4684/2021-01-27/CoRR4684ARetrospecti_DATA_2021-01-27_1430.csv", stringsAsFactors = F)
#can change "" into NA
noio <- readRDS("T:/Labs/Spakowicz/ppiio/data/curated/NoImmunotherapy.RDS")


P188 <- read.csv("T:/Labs/Spakowicz/ppiio/data/raw/P188/2021-01-27/P188NSCLCIOTox111611_DATA_2021-01-27_1441.csv")
comorb <- read_excel("T:/Labs/Spakowicz/ppiio/data/raw/CharlsonScore_Presley_08272019 acj.xlsx")
                                               
# T:/Labs/Spakowicz/projects/ppiio/data/raw/CoRR4684
# also read in the other db (P118)
# T:/Labs/Spakowicz/projects/ppiio/data/raw/P118

#[[[[these contain PHI -- no output to console that shows MRNS or dates, particularly if you're going to knit the Rmd]]]]

# safest to start by selecting only those variables you need e.g. age, gender, treatment, ppi, etc

```

# Format to cleaned ver

Note: No IO comes from a curated list of MRNs 
```{r}
# comorb.join <- comorb %>%
#   select(MRN, Total) %>%
#   rename("comorb.score" = "Total") %>%
#   mutate(MRN = as.character(MRN))
ppi_corr <- CoRR %>%
  rename("MRN"="record_id") %>%
#   filter(MRN %in% db$record_id) %>% 
#   select(MRN, dob, cancer, iostart, contains("stop"), io_2_start, io_3_start, vitalstatus, bmi,
#          ecog,sex,contains("ppi"), immunotherapy,chemo_prior,line_of_therapy, other_io, staging, deathdate,
#          lastcontact) %>%
#  select(-contains("subs"), -contains("other_ppi"), -contains("ppi_reason")) %>%
  filter(vitalstatus != 3) %>%
  mutate(vitalstatus = ifelse(vitalstatus == 1, 1, 0))%>%
  mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
  mutate(date = coalesce(deathdate, lastcontact),
         date = as.Date(date),
         iostart = as.Date(iostart),
         dob = as.Date(dob))%>%
  mutate(age=as.numeric(as.character(floor((date-dob)/365))))%>%
  filter(!is.na(iostart))%>%
  mutate(follow_up = as.numeric(as.character(date- iostart)))%>%
  left_join(cancer.key) %>%
#  (cancer.name %in% c("Other","melanoma","non-small cell lung cancer","small cell lung cancer"))%>%
# rename("comorb.score"="Total") %>%
  mutate(Cancer = if_else(grepl("melanoma|non-small cell lung cancer",
                                cancer.name),
                          true = cancer.name,
                          false = "Other"),
         Cancer = fct_relevel(Cancer, c("melanoma", "non-small cell lung cancer",
                                        "Other"))) %>%
  filter(immunotherapy != 11) %>%
  drop_na(iostart, ppi, date, dob) %>%
  filter(!MRN %in% noio)
 # left_join(comorb.join)

# table(ppi_corr$Cancer & )
   ##only throw out NA for survival info
  

 # mutate(deathdate = ifelse(deathdate == "", NA, deathdate),
 #         lastcontact = ifelse(lastcontact == "", NA, lastcontact)) %>%
 #  mutate(eventdate = coalesce(deathdate, lastcontact)) %>%
 #  mutate(eventdate = as.POSIXct(eventdate, format = "%F"),
 #         iostart = as.POSIXct(iostart, format = "%F")) %>%
 #  mutate(days = eventdate - iostart) 



############ no com.score ############
 P188$record_id <- as.character(P188$record_id)
 ppi_p188 <- P188 %>%
   rename("MRN"="record_id") %>%
   inner_join(comorb, by="MRN") %>%
   mutate(date_col = ifelse(deathdate == "", lastcontact,deathdate),
          dob = as.Date(dob),
          date_col = as.Date(date_col)) %>%
   filter(!is.na(date_col)) %>%
   mutate(age = floor(age_calc(dob,date_col, units = "years"))) %>%
   select(MRN,cancer,iostart, contains("stop"), -contains("subs"), io_2_start, io_3_start, vitalstatus, age, bmi, ecog,sex,
         contains("ppi"),-contains("other_ppi"), -contains("ppi_reason"), immunotherapy, chemo_prior,line_of_therapy, other_io, staging, deathdate, lastcontact,Total) %>%
   select(-c("chemo_stop","rad_stop")) %>%
   filter(!is.na(bmi) & !bmi == Inf) %>%
   filter(!is.na(immunotherapy)) %>%
   left_join(cancer.key)



 ppi_long <- rbind(ppi_p188, ppi_corr) %>%
   rename("comorb.score"="Total") %>%
   filter(!duplicated(MRN))
  

write.csv(ppi_corr, "../data/2021-01-27_db-formatted.csv", row.names = F)
```

summary
```{r}
# summary <- ppi_long %>%
#   group_by(cancer.name) %>%
#   summarise(n()) %>%
#   filter(`n()` > 3) %>%
#   filter(!cancer.name == "Other")
# 
# 
# # summary_tcc <- ppi_long_tcc %>%
# #   group_by(cancer.name) %>%
# #   summarise(n()) %>%
# #   filter(`n()` > 3) %>%
# #   filter(!cancer.name == "Other")

```


```{r}
# Create key for naming immunotherapy codes
ionames <- data.frame(immunotherapy = c(1,2,3,4,5,6,7,8,9,10), 
                      Immunotherapy = c("Nivolumab", "Pembrolizumab", 
                                        "Atezolizumab", "Ipilimumab", 
                                        "Nivolumab + Ipilimumab", 
                                        "Durvalumab + Tremelimumab", 
                                        "Tremelimumab", "Nivolumab + Chemotherapy",
                                        "Durvalumab", "Other"),
                      stringsAsFactors = FALSE)

# Select and adjust dataframe for Table 1
coxdf <- 
  ppi_corr %>%
  # select(MRN, vitalstatus, age, bmi, ecog, sex, cancer.name, contains("ppi"), 
  #       immunotherapy, chemo_prior, line_of_therapy,other_io, staging, comorb.score,follow_up) %>%
  # select(-contains("other_ppi"), -contains("ppi_reason")) %>%
  left_join(ionames) %>%
  mutate(`ECOG PS` = fct_relevel(ifelse(ecog > 2, ">2", ecog), "0","1","2",">2"),
         Staging  = ifelse(staging == 5, "Unknown", staging),
         BMI = bmi,
         Age = age,
         follow_up_time=follow_up,
         # Cancer = cancer.name, 
         # CCI = ifelse(comorb.score <= 1, "0-1", ">2"),
         Sex = sex,
         Therapy = ifelse(line_of_therapy == 1,"1",">1"),
         Chemo_prior = chemo_prior)
         # Immunotherapy = as.factor(Immunotherapy))
  # filter(!Immunotherapy == "Nivolumab + Chemotherapy")
         # Immunotherapy = ifelse(Immunotherapy == "Nivolumab + Chemotherapy",
         #                        "Other", Immunotherapy),
         # Immunotherapy = as.factor(Immunotherapy),
         # Immunotherapy = fct_relevel(Immunotherapy, "Atezolizumab", "Durvalumab",
         #                             "Durvalumab + Tremelimumab", "Ipilimumab", 
         #                             "Nivolumab", "Nivolumab + Ipilimumab", 
         #                             "Pembrolizumab", "Tremelimumab", "Other"))
```
 Adjust the immunotherapy to four categories
```{r}
coxdf <- 
  coxdf %>%
  # mutate(Immunotherapy = as.character(Immunotherapy)) %>%
  mutate(Immunotherapy = if_else(grepl("Atezo|Durvalumab$|Nivolumab$|Pembro",
                                           Immunotherapy),
                                     true = "PD1/L1",
                                     false = Immunotherapy),
         Immunotherapy = if_else(grepl("Durvalumab \\+ Tremelimumab$|Nivolumab \\+ Ipilimumab$",
                                           Immunotherapy),
                                     true = "PD1/L1 + CTLA4",
                                     false = Immunotherapy),
         Immunotherapy = if_else(grepl("Tremelimumab$|Ipilimumab$",
                                           Immunotherapy),
                                     true = "CTLA4",
                                     false = Immunotherapy),
         Immunotherapy = if_else(grepl("Nivolumab \\+ Chemotherapy",
                                           Immunotherapy),
                                     true = "Other",
                                     false = Immunotherapy),
         Immunotherapy = fct_relevel(Immunotherapy, c("PD1/L1", "CTLA4", "PD1/L1 + CTLA4", "Other")))
table(coxdf$Immunotherapy)
table(coxdf$Immunotherapy)
         
```


##CHECK MEDICATION; check p val for different catagory between CCI/ECOG &ppi
# Creating table 1

```{r}
listVars <- c("Age", "Sex", "vitalstatus","ECOG PS","Cancer", "Staging", "Immunotherapy","Chemo_prior","follow_up_time")
catVars <- c("Sex", "Cancer", "vitalstatus","Immunotherapy", "ECOG PS","Staging")
table1 <- CreateTableOne(vars = listVars,
                         data = coxdf,
                         factorVars = catVars,
                         strata = c("ppi","Therapy"))
tbl1.ppi.therapy <- print(table1, nonnormal = "follow_up_time")
######################################################################################
listVars2 <- c("Age", "Sex", "vitalstatus","ECOG PS","Cancer", "Staging", "Immunotherapy","Therapy","ppi","follow_up_time")
catVars2 <- c("Sex", "Cancer", "vitalstatus","Immunotherapy", "ECOG PS","Staging","Therapy","ppi")
table2 <- CreateTableOne(vars = listVars2,
                         data = coxdf,
                         factorVars = catVars2,
                         strata = c("ppi", "Chemo_prior"))
tbl1.ppi.chemoprior <- print(table2, nonnormal = "follow_up_time")
######################################################################################
listVars3 <- c("Age", "Sex", "vitalstatus","ECOG PS","Cancer", "Staging", "Immunotherapy","Therapy","Chemo_prior","follow_up_time")
catVars3 <- c("Sex", "Cancer","vitalstatus", "Immunotherapy", "ECOG PS","Therapy","Staging")
table3 <- CreateTableOne(vars = listVars3,
                         data = coxdf,
                         factorVars = catVars3,
                         strata = "ppi")
tbl1.ppi <- print(table3, nonnormal = "follow_up_time")
######################################################################################
```

Write table versions to files.
```{r write}
write.csv(tbl1.ppi.therapy, "../tables/table1_ppi-lotherapy.csv")
write.csv(tbl1.ppi.chemoprior, "../tables/table1_ppi-chemoprior.csv")
write.csv(tbl1.ppi, "../tables/table1_ppi.csv")
```


```{r}
COMPARE <- coxdf %>%
  group_by(CCI, ppi) %>%
  tally()

ggplot(coxdf,aes(x=CCI, y=ppi, fill = ppi))+
  geom_bar(stat="identity")


```

What I need:
+ PPI stop dates
+ Surgical collection dates (look at the scratch sheet)
  + time from collection to last follow up or death (it is in
+ How to connect ppi start and stop dates by a line
+ Add second ppi use date



```{r}
x <- CoRR%>%
  select(record_id,bmi) %>%
  rename("MRN"="record_id") %>%
  inner_join(comorb, by="MRN")

y <- P188%>%
select(record_id,bmi) %>%
  rename("MRN"="record_id") %>%
  inner_join(comorb, by="MRN")

# flow_comp %>%
#   ggplot(aes(x = days.from.start, y = subject)) +
#   # Samples colored based on response
#   geom_point(aes(color = response, shape = arm))




# ppi_long %>%
#     ggplot() +
#     geom_point(aes(x = ppidays.to.iostart, y = as.numeric(as.factor(record_id)), color = ppidays.to.iostart <= 0)) +
#     geom_point(aes(x = days.to.ioend, y = as.numeric(as.factor(record_id)))) +
#     labs(title="PPI and IO Timeline for Each Patient",
#         y ="Patient", x = "PPI and IO Timeline (IO start = 0")
#     
# 
# ppi_long %>%
#   filter(ppidays.to.iostart >= -60) %>%
#     ggplot() +
#     geom_point(aes(x = ppidays.to.iostart, y = as.numeric(as.factor(record_id)), color = ppidays.to.iostart <= 0)) +
#     geom_point(aes(x = days.to.ioend, y = as.numeric(as.factor(record_id)))) +
#     labs(title="PPI and IO Timeline for Each Patient \n (filtered to PPIs 60 days before IO start)",
#         y ="Patient", x = "PPI and IO Timeline (IO start = 0")
# 
# 
# ppi_long_tcc %>%
#     ggplot() +
#     geom_point(aes(x = ppidays.to.iostart, y = as.numeric(as.factor(record_id)), color = ppidays.to.iostart <= 0)) +
#     geom_point(aes(x = days.to.ioend, y = as.numeric(as.factor(record_id)))) +
#     labs(title="PPI and IO Timeline for Each Patient with TCC Data",
#         y ="Patient", x = "PPI and IO Timeline (IO start = 0")
# 
# ppi_long_tcc %>%
#   filter(ppidays.to.iostart >= -365) %>%
#     ggplot() +
#     geom_point(aes(x = ppidays.to.iostart, y = as.numeric(as.factor(record_id)), color = ppidays.to.iostart <= 0)) +
#     geom_point(aes(x = days.to.ioend, y = as.numeric(as.factor(record_id)))) +
#     labs(title="PPI and IO Timeline for Each Patient with TCC Data \n (filtered to PPIs 365 days before IO start)",
#         y ="Patient", x = "PPI and IO Timeline (IO start = 0")


```


```{r}
orien <- read.csv("T:/Labs/Spakowicz/projects/ppiio/data/derived/2021-03-23/TCC000138_formatted.csv",
                  stringsAsFactors = F)
```

```{r}
table(orien$PPI_collection, orien$H2B_collection)

table(orien$PPI_collection)
```


