---
title: "Check FASTQ checksums (Cohort 3 / Iobio)"
author: "Dan Spakowicz"
date: "`r Sys.Date()`"
output: html_document
---

Verifies that the 14 paired FASTQ libraries copied to the lab share match the
sequencing-core md5 manifest. The OSC archive of the same files is
`/fs/ess/PAS1695/archive/projects/iobio/` (see
`processing_prepare-humann3-batches.R`).

Sequencer filenames include collection dates. Map them with
`Iobio_sample_map` from `00-paths.R` (CSV on the share, not in git). The RDS
written below stores **study IDs** (`P026_d0_R1.fastq.gz`, …) and checksums
only — not share paths.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tools)
library(magrittr)
library(dplyr)

this_dir <- if (!is.null(knitr::current_input())) {
  dirname(normalizePath(knitr::current_input()))
} else {
  getwd()
}
paths_file <- file.path(this_dir, "..", "..", "manuscript", "scripts", "00-paths.R")
if (!file.exists(paths_file)) {
  stop("Missing manuscript/scripts/00-paths.R (needed for Tdrive / Iobio_sample_map).")
}
source(paths_file)
```

```{r}
files <- list.files(path = "T:/Labs/Spakowicz/Iobio/data/raw/fastqs/")

files
```

```{r}
key_candidates <- c(
  if (exists("Iobio_sample_map")) Iobio_sample_map,
  if (exists("Tdrive")) file.path(Tdrive, "iobio-sample-key.csv")
)
key_file <- key_candidates[!is.na(key_candidates) & file.exists(key_candidates)][1]
if (is.na(key_file)) {
  stop("Missing iobio-sample-key.csv. Keep it on the lab share; do not commit it.")
}
sample_map <- read.csv(key_file, stringsAsFactors = FALSE)

files.full <- list.files(path = "T:/Labs/Spakowicz/Iobio/data/raw/fastqs/",
                         full.names = TRUE)

x <- read.table(files.full[grep("md5", files.full)]) %>%
  rename("md5sum" = "V1",
         "file" = "V2") %>%
  mutate(new.loc = files.full[grep("gz", files.full)])

# Calculate md5sums (takes a long time)
new.sum <-
  x %>%
  mutate(new.loc.md5 = md5sum(new.loc))

checked <-
  new.sum %>%
  mutate(TransferredCorrectly = md5sum == new.loc.md5) %>%
  mutate(
    sequencer = sub("_R([12])_001\\.fastq\\.gz$", "", basename(file)),
    read = sub(".*_R([12])_001\\.fastq\\.gz$", "R\\1", basename(file))
  ) %>%
  left_join(sample_map, by = "sequencer") %>%
  transmute(
    file = paste0(sample, "_", read, ".fastq.gz"),
    md5sum,
    copy_md5 = new.loc.md5,
    TransferredCorrectly
  )

saveRDS(checked, file = "../data/md5sums-ck.RDS")
```
